### 2019-03-07 Paper_[2019-02]_Numerical_Algorithms/paper.tex

\RequirePackage{fix-cm}
\documentclass[smallextended,final]{svjour3}       % onecolumn (second format)
\smartqed  % flush right qed marks, e.g. at end of proof
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{tikz}     % For adding axes to plots
\usetikzlibrary{positioning}
\usepackage{graphicx} % For including PDFs
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{booktabs} % For formal tables
\usepackage[colorlinks,allcolors=blue]{hyperref}

%
% please place your own definitions here and don't use \def but
% \newcommand{}{}

%% Command for creating multiplie-line cells inside a table environment
\newcommand{\multilinecell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

%% Create a ``Lemma'' environment that does not have italics font.
\spnewtheorem{plainlemma}{Lemma}{\bf}{}
%% Create a ``Theorem'' environment that is not numbered.
\spnewtheorem*{onetheorem}{Theorem}{\bf}{}
\spnewtheorem*{proofdot}{Proof.}{\it}{}

%% \newcommand{\Bigl}{\mathopen\Big}
%% \newcommand{\Bigr}{\mathclose\Big}
%
% Insert the name of "your journal" with
\journalname{Numerical Algorithms}
%
\begin{document}

\title{Interpolation of Sparse High-Dimensional Data\thanks
  {This work was supported by the National Science Foundation Grant CNS-1565314.}
}
% \subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Thomas C. H. Lux   \and
  Layne T. Watson          \and
  Tyler H. Chang           \and
  Yili Hong                \and
  Kirk Cameron
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Thomas Lux \at
  Virginia Polytechnic Institute \& State University (VPI\&SU) \\
  Blacksburg, VA 24061 \\
  \email{tchlux@vt.edu}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
A rapid increase in the quantity of data available is allowing all
fields of science to generate more accurate models of multivariate
phenomena. Regression and interpolation become challenging when the
dimension of data is large, especially while maintaining tractable
computational complexity. While regression is a popular approach to
solving approximation problems with high dimension, there are often
advantages to interpolation. This paper presents error bounds for
interpolants on moderately high dimensional problems and contrasts
their performance with that of popular regression
techniques. Empirical results demonstrate the viability of
interpolants for high dimensional approximation problems, suggesting
that these techniques are capable of effectively modeling multivariate
phenomena while maintaining flexibility in different application
domains.

\keywords{approximation \and regression \and interpolation \and high
  dimension \and error bound}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

%     Introduction     
%======================

%% ===================================================================
\section{Introduction}
\label{sec:introduction}

Regression and interpolation are problems of considerable importance
that find applications across many fields of science. Pollution and
air quality analysis \cite{de2008field}, energy consumption management
\cite{lazos2014optimisation}, and student performance prediction
\cite{cortez2008using,lux2016applications} are a few of many
interdisciplinary applications of multivariate regression for
predictive analysis. As discussed later, these techniques can also be
applied to prediction problems related to forest fire risk assessment
\cite{cortez2007data}, Parkinson's patient clinical evaluations
\cite{tsanas2010accurate}, local rainfall and weather
\cite{williams2009rattle}, credit card transactions
\cite{pozzolo2015calibrating}, and high performance computing (HPC)
file input/output (I/O) \cite{lux2018nonparametric}.

Regression and interpolation have a considerable theoretical base in
one dimension \cite{cheney2009course}. Splines in particular are well
understood as an interpolation technique in one dimension
\cite{de1978practical}, particularly B-splines. Tensor products of
B-splines \cite{unther1996interpolating} or other basis functions have
an unfortunate exponential scaling with increasing
dimension. Exponential scaling prohibits tensor products from being
reasonably applied beyond three-dimensional data. In order to address
this dimensional scaling challenge, C. de Boor and others proposed box
splines \cite{de2013box}, of which one of the approximation techniques
in this work is composed \cite{lux2018novel}.

The theoretical foundation of low dimensional interpolation allows the
construction of strong error bounds that are absent from high
dimensional problems. This work extends some known results regarding
the secant method \cite{dennis1996numerical} to construct an
interpolation error bound for problems of arbitrary dimension. These
error bounds are useful, considering the same cannot be said for
regression algorithms in general. The maximum complexity of an
interpolant is bounded by the amount of data available, while the
maximum complexity of a regressor is bounded by both the amount of
data and the chosen parametric form. For this reason, generic uniform
bounds are largely unobtainable for regression techniques on arbitrary
approximation problems, even when the approximation domain is bounded.

Aside from theoretical motivation for the use of interpolants, there
are often computational advantages as well. Interpolants do not have
the need for \textit{fitting} data, or minimizing error with respect
to model parameters. In applications where the amount of data is large
and the relative number of predictions that need to be made for a
given collection of data is small, the direct application of an
interpolant is much less computationally expensive.

In this work, multivariate interpolation is defined given a closed
convex subset $Y$ of a metrizable topological vector space with metric
$s$, some function $f:\mathbb{R}^d \rightarrow Y$ and a set of points
$X = \bigl\{x^{(1)}$,$\ldots$,$x^{(n)}\bigr\} \subset \mathbb{R}^d,$
along with associated function values $f\bigl(x^{(i)}\bigr)$. The goal
is to construct an approximation $\hat f: \mathbb{R}^d \rightarrow Y$
such that $\hat f\bigl(x^{(i)}\bigr) = f\bigl(x^{(i)}\bigr)$ for all
$i = 1$,$\ldots$,$n$. It is often the case that the form of the true
underlying function $f$ is unknown, however it is still desirable to
construct an approximation $\hat f$ with small approximation error at
$y \notin X$. The two metric spaces that will be discussed in this
work are the real numbers with metric $s(x,y) = |x-y|$, and the set of
cumulative distribution functions (CDFs) with the Kolmogorov-Smirnov
(KS) statistic as metric.

Multivariate regression is often used when the underlying function is
presumed to be stochastic, or stochastic error is introduced in the
evaluation of $f$. Hence, multivariate regression relaxes the
conditions of interpolation by choosing parameters $P$ defining $\hat
f(x;P)$ to minimize the error vector $\Bigl( \bigl | \hat f
\bigl(x^{(1)};P\bigr) - f\bigl(x^{(1)}\bigr) \bigr|$, $\ldots$, $\bigl
| \hat f \bigl(x^{(n)}; P\bigr) - f\bigl(x^{(n)}\bigr) \bigr | \Bigr)$
in some norm. The difficult question in the case of regression is
often what parametric form to adopt for any given application.

The most challenging problem when scaling in dimension is that the
number of possible interactions between dimensions grows
exponentially. Quantifying all possible interactions becomes
intractable and hence beyond three-dimensional data, mostly linear
models are used. That is not to say nonlinear models are absent, but
nonlinearities are often either preconceived or model pairwise
interactions between dimensions at most. Even globally nonlinear
approximations such as neural networks are constructed from
compositions of summed low-interaction functions
\cite{clevert2015fast}.

Provided the theoretical and practical motivations for exploring
interpolants, this work aims to study the empirical performance
differences between a set of scalable (moderately) interpolation
techniques and a set of common regression techniques. These techniques
are applied to a collection of moderately high dimensional problems
($5 \le d \le 30$) and the empirical results are discussed.

The remainder of this paper is organized as follows. Sections
\ref{sec:regression} and \ref{sec:interpolation} present the
multivariate models. Section \ref{sec:error} presents the error
measuring methodology that is used for collecting empirical
results. Section \ref{sec:theory} presents the theoretical error
bounds for interpolation. Section \ref{sec:data} presents the data
sets for empirical approximation analysis and presents the
approximation results for all models on these data sets. Section
\ref{sec:discussion} analyzes and discusses the results, their
implications, and their limitations. Finally, Section
\ref{sec:conclusion} concludes.

% Art is a subjective form of personal expression and computers are
% just robots that are going to take over the world. \textit{*deep
% breath*} In order to communicate my frustration with the
% technological and information revolution, I'm going to write the
% rest of this paper in poetry.

% Star light, star bright \ldots

%% ===================================================================
\section{Multivariate Regression}
\label{sec:regression}
Multivariate regressors are capable of accurately modeling a complex
dependence of a response (in $Y$) on multiple variables (represented
as a points in $\mathbb{R}^{d}$). The approximations to some (unknown)
underlying function $f: \mathbb{R}^d \rightarrow Y$ are chosen to
minimize some error measure related to data samples
$f\bigl(x^{(i)}\bigr)$. For example, least squares regression uses the
sum of squared differences between modeled function values and true
function values as an error measure. In this section and the next,
some techniques are limited to approximating real valued functions ($Y
\subset \mathbb{R}$). These techniques can be extended to real
vector-valued ranges by repeating the construction for each component
of the vector output. Throughout the following, $x$ denotes a
$d$-tuple, $x_i$ the $i$th component of $x$, and $x^{(i)}$ the $i$th
$d$-tuple data point. Different symbols are used to represent the
approximation function $\hat f$.

\subsection{Multivariate Adaptive Regression Splines}
\label{sec:mars}
This approximation was introduced in \cite{friedman1991multivariate}
and subsequently improved to its current version in
\cite{stanford1993fast}, called fast multivariate adaptive regression
splines (Fast MARS).In Fast MARS, a least squares fit model is
iteratively built by beginning with a single constant valued function
and adding two new basis functions at each iteration of the form
\begin{align*}
  B_{2j-1}(x) &= B_l(x) \bigl(x_i-x^{(p)}_i\bigr)_+, \\
  B_{2j}(x) &= B_k(x) \bigl(x_i-x^{(p)}_i\bigr)_- ,
\end{align*}
where $j$ is the iteration number, $1 \le p \le n$, and $B_l(x)$,
$B_k(x)$ are basis functions from the previous iteration,
$$w_+ = \begin{cases} w, & w \geq 0 \\ 0, & w < 0 \end{cases},$$
and $w_- = (-w)_+$. After iteratively constructing a model, MARS then
iteratively removes basis functions that do not contribute to goodness
of fit. In effect, MARS creates a locally component-wise tensor
product approximation of the data. The overall computational
complexity of Fast MARS is $\mathcal{O}(n d m^3)$ where $m$ is the
maximum number of underlying basis functions. This paper uses an
implementation of Fast MARS \cite{rudy2017pyearth} with $m = 200$
throughout all experiments.

\subsection{Support Vector Regressor}
\label{sec:svr}
Support vector machines are a common method used in machine learning
classification tasks that can be adapted for the purpose of regression
\cite{basak2007support}. How to build a support vector regressor (SVR)
is beyond the scope of this summary, but the resulting functional fit
$p : \mathbb{R}^d \rightarrow \mathbb{R}$ has the form
$$ p(x)  = \sum_{i=1}^{n}a_i K\bigl(x,x^{(i)}\bigr) + b ,$$
where $K$ is the selected kernel function, $a_i \in \mathbb{R}^n$, 
$b \in \mathbb{R}$ are coefficients to be solved for simultaneously. The
computational complexity of the SVR is $\mathcal{O}(n^2dm)$, with $m$
being determined by the minimization convergence criterion. This paper
uses the scikit-learn SVR \cite{scikit-learn} with a polynomial kernel
function.

\subsection{Multilayer Perceptron Regressor}
\label{sec:mlp}
The neural network is a well studied and widely used method for both
regression and classification tasks
\cite{rumelhart1988learning,hornik1989multilayer}. When using the
rectified linear unit (ReLU) activation function
\cite{dahl2013improving} and fitting the model with stochastic
gradient descent (SGD) or BFGS minimization techniques
\cite{goh2017why,moller1993scaled,robbins1951stochastic}, the model
built by a multilayer perceptron uses layers $l : \mathbb{R}^{i}
\rightarrow \mathbb{R}^{j}$ defined by
$$ l(u) = \bigl( u^t W_l \bigr)_+ ,$$
where $W_l$ is the $i$ by $j$ weight matrix for layer $l$. In this
form, the multilayer perceptron (MLP) produces a piecewise linear
model of the input data. The computational complexity of training a
multilayer perceptron is $\mathcal{O}(n d m)$, where $m$ is determined
by the sizes of the layers of the network and the stopping criterion
of the minimization used for finding weights. This paper uses a MLP
built from Keras and Tensorflow to perform regression
\cite{chollet2015keras,tensorflow2015-whitepaper} with ten hidden
layers each having thirty two nodes (a total of approximately ten
thousand parameters), ReLU activation, and five thousand epochs of SGD
for training.

%% ===================================================================
\section{Multivariate Interpolation}
\label{sec:interpolation}
The following interpolation techniques demonstrate a reasonable
variety of approaches to interpolation. All of the presented
interpolants produce approximations that are continuous in value,
which is often a desirable property in applied approximation problems.

\subsection{Delaunay}
\label{sec:delaunay}

The Delaunay triangulation is a well-studied geometric technique for
producing an interpolant \cite{lee1980two}. The Delaunay triangulation
of a set of data points into simplices is such that there are no data
points inside the sphere defined by the vertices of each simplex. For
a $d$-simplex S with vertices $x^{(0)}$, $x^{(1)}$, $\ldots$,
$x^{(d)}$, and function values $f\bigl(x^{(i)}\bigr)$, $i=0$,
$\ldots$, $d$, $y \in S$ is a unique convex combination of the
vertices,
$$ y = \sum_{i=0}^{d} w_i x^{(i)}, \quad \sum_{i=0}^{d} w_i = 1, \quad w_i \geq 0, \quad i=0,\ldots,d, $$
and the Delaunay interpolant $\hat f(y)$ at $y$ is given by
$$ \hat f(y) = \sum_{i=0}^{d} w_i f\bigl(x^{(i)}\bigr). $$
The computational complexity of Delaunay interpolation (for the
implementation used \cite{chang2018polynomial}) is $\mathcal{O}(n d^4
\log d)$, which is capable of scaling reasonably to $d \leq 50$. In
the present application, a Delaunay simplex $S$ containing $y$ is
found, then the $d+1$ vertices (points in $X$) of $S$ are used to
assign weights to each vertex and produce the predicted function value
for point $y$.

\subsection{Modified Shepard}
\label{sec:modified-shepard}

The modified Shepard method used here (ShepMod) generates a continuous
approximation based on Euclidean distance and resembles a nearest
neighbor interpolant \cite{cover1967nearest}. This model is a type of
\textit{metric interpolation}, also called a Shepard method
\cite{shepard1968two,gordon1978shepard}. The interpolant has the form
$$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)f\bigl(x^{(k)}\bigr)}{\sum\limits_{k=1}^{n}W_k(x)},$$
where $x \not = x^{(k)}$, $k = 1$, $\ldots$, $n$, for weighting functions
$$W_k(x) = \left( \frac{\big(r_k - \bigl\|x - x^{(k)}\bigr\|_2\big)_+}{r_k \bigl\|x - x^{(k)}\bigr\|_2} \right)^2,$$ 
where $r_k \in \mathbb{R}$ are chosen such that at least $d+1$ other
points $x^{(j)}$, $j \not = k$ are inside the closed Euclidean ball of
radius $r_k$ about $x^{(k)}$. The computational complexity of ShepMod
is $\mathcal{O}(n^2d)$. This paper uses a Fortran 95 implementation of
ShepMod derived from SHEPPACK \cite{thacker2010algorithm}.

\subsection{Linear Shepard}
The linear Shepard method (LSHEP) is a blending function using local
linear interpolants, a special case of the general Shepard algorithm
\cite{thacker2010algorithm}. The interpolant has the form
$$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)P_k(x)}{\sum\limits_{k=1}^{n}W_k(x)} ,$$
where $W_k(x)$ is a locally supported weighting function and $P_k(x)$
is a local linear approximation to the data satisfying
$P_k\bigl(x^{(k)}\bigr) = f\bigl(x^{(x)}\bigr)$. The computational
complexity of LSHEP is $\mathcal{O}(n^2d^3)$. This paper uses the
Fortran 95 implementation of LSHEP in SHEPPACK
\cite{thacker2010algorithm}.

\subsection{Box Splines}
\label{sec:box-splines}

The box spline model used here is an interpolation technique built
from overlapping box splines \cite{de2013box}. The box splines serve
as basis functions that are shifted and scaled to have support over
box shaped regions. The boxes are constructed in a way to guarantee a
covering of the domain \cite{lux2018novel}. Given a set of box splines
$\bigl\{b^{x^{(i)}}\bigr\}$ with the iterative box properties outlined
in \cite{lux2018novel} and anchored at interior points
$\bigl\{x^{(i)}\bigr\}$,
$$ \hat f(y) = \frac{\sum\limits_{i=1}^n b^{x^{(i)}}(y) f\bigl(x^{(i)}\bigr)}{\sum\limits_{i=1}^n b^{x^{(i)}}(y)}. $$
Note that the box splines always satisfy
$b^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $b^{x^{(i)}}(y)
\geq 0$. The computational complexity of interpolation via the box
spline model is $\mathcal{O}(n^2 d)$.

\subsection{Voronoi}
\label{sec:voronoi}

A well-studied technique for classification and approximation is the
nearest neighbor algorithm \cite{cover1967nearest}. Nearest neighbor
inherently utilizes the convex region $C^{x^{(i)}}$ (Voronoi cell
\cite{dirichlet1850reduction}) consisting of all points closer to
$x^{(i)}$ than to any other point $x^{(j)}$. The Voronoi model smooths
the nearest neighbor approximation by utilizing the Voronoi cells to
define support via a generic basis function $v: \mathbb{R}^d
\rightarrow \mathbb{R}_+$ given by
$$ v^{x^{(i)}}(y) = \left(1 - \frac{\bigl\|y - x^{(i)}\bigr\|_2}{2 \ h\bigl(y - x^{(i)} \mid x^{(i)}\bigr)} \right)_+, $$
where $h\bigl(w \mid x^{(i)}\bigr)$ is the distance between $x^{(i)}$
and the boundary of the Voronoi cell $C^{x^{(i)}}$ in the direction
$w$. $v^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $v^{x^{(i)}}$
has local support, giving the interpolated value
$$ f(y) = \frac{\sum\limits_{i=1}^n v^{x^{(i)}}(y) f\bigl(x^{(i)}\bigr)}{\sum\limits_{i=1}^n v^{x^{(i)}}(y)}, $$
where $0 \leq v^{x^{(i)}}(y) \leq 1$. The computational complexity of
interpolation via this Voronoi model is $\mathcal{O}(n^2 d)$. All of
the approximations are an interpolant involving a convex combination
of known function values $f\bigl(x^{(i)}\bigr)$.

%% ===================================================================
\section{Measuring Error}
\label{sec:error}

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/plot-delaunay-example-prediction.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Throughput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {CDF Value};
  \end{tikzpicture}
  \vspace{-0.3cm}
  \caption{In this HPC I/O example, the general methodology for
    predicting a CDF and evaluating error can be seen. The Delaunay
    method chose three source distributions (dotted lines) and
    assigned weights \{.3, .4, .3\} (top to bottom at arrow). The
    weighted sum of the three known CDFs produces the predicted CDF
    (dashed line). The KS Statistic (arrow) computed between the true
    CDF (solid line) and predicted CDF (dashed line) is 0.2 for this
    example. The KS test null hypothesis is rejected at $p$-value
    0.01, however it is not rejected at $p$-value 0.001.
  \vspace{-.1cm}}
  \label{fig:prediction-example}
\end{figure}

When the range of an approximation is the real numbers, error is
reported with summary statistics including: min absolute error, max
absolute error, and absolute error quartiles. When the range of an
approximation is the space of cumulative distribution functions, the
Kolmogorov-Smirnov statistic (max-norm difference between the
functions) is used.

A hurdle when modeling function-valued outputs such as cumulative
distribution functions (CDFs) or probability density functions (PDFs)
is that certain properties must be maintained. It is necessary that a
PDF $f: \mathbb{R} \rightarrow \mathbb{R}$ have the properties $f(x)
\geq 0$ and $\int_{-\infty}^{\infty}f(x)dx = 1$. However, for a CDF
$F: \mathbb{R} \rightarrow \mathbb{R}$ the properties are
$F(x) \in [0,1]$ and $F(x)$ is absolutely continuous and
nondecreasing. This work utilizes the fact that a convex combination
of CDFs (or PDFs) results in a valid CDF. Given $G(x) = \sum_{i}w_i
F_i(x)$, $\sum_{i} w_i = 1$, $w_i \geq 0$, and each $F_i$ is a valid
CDF, $G$ must also be a valid CDF. A demonstration of how this is
applied can be seen in Figure \ref{fig:prediction-example}.

The performance of approximation techniques that predict probability
functions can be analyzed through a variety of summary
statistics. This work uses the max absolute difference, also known as
the Kolmogorov-Smirnov (KS) statistic \cite{lilliefors1967kolmogorov}
for its compatibility with the KS test.

The two-sample KS test is a useful nonparametric test for comparing
two CDFs while only assuming stationarity, finite mean, and finite
variance. The null hypothesis (that two CDFs come from the same
underlying distribution) is rejected at level $p \in [0,1]$ when
$$ KS > \sqrt{-\frac{1}{2}\ln\biggl(\frac{p}{2}\biggr)} \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}, $$
with distribution sample sizes $n_1,n_2 \in \mathcal{N}$. For all
applications of the KS test presented in this work $n_1 = n_2$. An
example of the round-trip prediction methodology from known and
predicted distributions to the calculation of error can be seen in
Figure \ref{fig:prediction-example}.

\section{Theoretical Error Bound}
\label{sec:theory}

This section presents the theoretical results bounding the error of
simplicial interpolation. The error analysis relies on simplicial
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly
apply to Delaunay; and (3) multiple other interpolants in this paper
compute predictions as convex combinations of observed function
values, which may allow for straight forward extensions of this error
bound.

\begin{plainlemma}
  \label{lemma:1}
  Let $S \subset \mathbb{R}^d$ be open and convex, $f: \mathbb{R}^d
  \rightarrow \mathbb{R}$, and $\nabla f \in
  Lip_{(\gamma,\|\cdot\|_2)}(S)$, the set of $\gamma$-Lipschitz
  continuous functions in the $2$-norm. Then for all $x,y \in S$
  $$\big|f(y) - f(x) - \langle \nabla f(x), y - x \rangle \big| \leq \frac{\gamma \|y - x\|_2^2}{2}.$$
\end{plainlemma}

\begin{proofdot}
  Consider the function $g(t) = f \big((1-t) x + t y \big)$, $0 \leq t
  \leq 1$, whose derivative $g'(t) = \big\langle \nabla f \big((1-t) x
  + t y \big), y - x \big\rangle$ is the directional derivative of $f$
  in the direction $(y - x).$
  \begin{align*}
    \big|f(y) - f(x) - \langle \nabla f(x), y - x \rangle \big|
       &= \big|g(1) - g(0) - g'(0) \big| & \\
       &= \bigg| \int_0^1 g'(t) - g'(0)\ dt \bigg| \\
       &\leq \int_0^1 \big|g'(t) - g'(0)\big|\ dt \\
       &= \int_0^1 \bigg| \big \langle \nabla f\big((1-t)x + ty\big) - \nabla f(x), y - x \big \rangle \bigg|\ dt \\
       &\leq \int_0^1 \big \| \nabla f\big((1-t)x + ty\big) - \nabla f(x) \big \|_2\ \| y - x \|_2\ dt \\
       &\leq \int_0^1 \big ( \gamma\ \|y-x\|_2 \big) \ \big( \|y-x\|_2 \big) t\ dt \\
       &= \frac{\gamma \|y - x\|_2^2}{2}.
  \end{align*}
  \qed
\end{proofdot}

\begin{plainlemma}
  \label{lemma:2}
  Let $x, y, v_i \in \mathbb{R}^d$, $c_i \in \mathbb{R}$, and
  $|\langle y - x, v_i \rangle| \leq c_i$ for $i = 1$, $\ldots$, $d.$
  If $M = (v_1$, $\ldots$, $v_d)$ is nonsingular, then
  $$\|y - x\|_2^2 \leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2,$$
  where $\sigma_d$ is the smallest singular value of $M.$
\end{plainlemma}

\begin{proofdot}
  Using the facts that $M$ and $M^t$ have the same singular values,
  and $\|M^tw\|_2 \geq \sigma_d \|w\|_2$, gives
  \begin{align*}
    \|y - x\|_2^2 &\leq \frac{\|M^t (y - x)\|_2^2}{\sigma_d^2} \\
                  &=    \frac{1}{\sigma_d^2} \sum_{i=1}^d \langle y - x, v_i \rangle^2 \\
                  &\leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2.
  \end{align*}
  \qed
\end{proofdot}

\begin{plainlemma}
  \label{lemma:3}
  Given $f$, $\gamma$, $S$ as in {\it Lemma \ref{lemma:1}}, let $X =
  \{x_0$, $x_1$, $\ldots$, $x_d\}$ $\subset S$ be the vertices of a
  $d$-simplex, and let $\hat f(x) = \langle c, x - x_0 \rangle +
  f(x_0)$, $c \in \mathbb{R}^d$ be the linear function interpolating
  $f$ on $X.$ Let $\sigma_d$ be the smallest singular value of the
  matrix $M = (x_1 - x_0$, $\ldots$, $x_d - x_0)$, and $k =
  \max\limits_{1\ \leq\ j\ \leq\ d} \|x_j - x_0\|_2.$ Then
  $$\big\|\nabla f(x_0) - c\big\|_2 \leq \sqrt{d} \frac{\gamma k^2}{\sigma_d}.$$
\end{plainlemma}

\begin{proofdot}
  Consider $f(x) - \hat f(x)$ along the line segment $z(t) = (1-t)x_0
  + t x_j$, $0 \leq t \leq 1.$ By Rolle's Theorem, for some $0 < \hat
  t < 1$, $\big\langle \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0
  \big\rangle = 0.$ Now
  \begin{align*}
    \big| \big< \nabla f(x_0) - c, x_j - x_0 \big \rangle \big|
        &= \big| \big< \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) + \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0 \big \rangle \big| \\
	&= \big| \big\langle \nabla f(x_0) - \nabla f \big(z(\ \hat t\ )\big), x_j - x_0 \big \rangle \big| \\
	&\leq \big \| \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) \big \|_2 \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_0 - z(\ \hat t\ )\|_2\ \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_j - x_0\|_2^2 \leq \gamma k^2, \\
  \end{align*}
  for all $1 \leq j \leq d.$ Using {\it Lemma \ref{lemma:2}},
  $$\big \| \nabla f(x_i) - c \big \|_2^2 \leq \frac{d}{\sigma_d^2}
  \big( \gamma k^2\big)^2 \implies \big \| \nabla f(x_i) - c \big \|_2
  \leq \sqrt{d} \frac{\gamma k^2}{\sigma_d}.$$ \qed
\end{proofdot}

\begin{onetheorem}
  Under the assumptions of {\it Lemma \ref{lemma:1}} and {\it Lemma \ref{lemma:3}}, for $z \in S$,
  $$ \big|f(z) - \hat f(z)\big| \leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d} \frac{\gamma k^2}{\sigma_d} \|z - x_0\|_2.$$
\end{onetheorem}

\begin{proofdot}
  Let $v = \nabla f(x_0) - c$, where $\|v\|_2 \leq \sqrt{d} \gamma k^2 / \sigma_d$ by {\it Lemma \ref{lemma:3}}. Now
  \begin{align*}
    \big|f(z) - \hat f(z)\big|
       &= \big|f(z) - f(x_0) - \langle c, z - x_0 \rangle \big| \\ 
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) - v, z - x_0 \rangle \big| \\
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle + \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \big| \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \|v\|_2 \|z - x_0\|_2 \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0), z - x_0 \rangle \big| + \textstyle{\frac{\gamma k^2 \sqrt{d}}{\sigma_d}} \|z - x_0\|_2 \\
       &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d}\frac{\gamma k^2}{\sigma_d} \|z - x_0\|_2,
  \end{align*}
  where the last inequality follows from {\it Lemma \ref{lemma:1}}.
  %
  \qed
\end{proofdot}

In summary, the approximation error of a simplicial interpolant tends
quadratically towards zero when approaching observed data only when
the diameter of the simplex is also reduced at a proportional rate.
Only linear convergence to the true function can be achieved in
practice, without the incorporation of additional observations. Notice
that the approximation error is largely determined by the spacing of
observed data. Predictions made by simplices whose vertices are not
well-spaced (i.e. have large diameter, or are nearly contained in a
hyperplane) have higher error. In light of this error bound, an
empirical evaluation of presented algorithms follows.

%% ===================================================================
\section{Data and Empirical Analysis}
\label{sec:data}

The theoretical results constructed in Section \ref{sec:theory} for
simplicial interpolation are promising, however they are difficult to
interpret in context with approximation algorithms that do not have
similar known uniform error bounds. For that reason, this paper
utilizes five different data sets of varying dimension and application
to construct approximations and compare the accuracy of different
techniques.

In the following five subsections the sources and targets of each test
data set are described, as well as challenges and limitations related
to approximating these data. The distribution of response values being
modeled is presented followed by the distribution of approximation
errors for each algorithm.

All five data sets are rescaled such that the domain of approximation
is the unit hypercube. The range of the first four data sets is the
real numbers, while the range of the fifth data set is the space of
cumulative distribution functions. All approximation techniques are
applied to the first four data sets, while only those interpolants
whose approximations are convex combinations of observed data are
applied to the final data set.

All approximations are constructed using $k$-fold cross validation as
described in \cite{kohavi1995study} with $k=10$. This approach
randomly partitions data into $k$ nearly equal sized sets. Each
algorithm is then evaluated by constructing an approximation over each
unique union of $k-1$ elements of the partition, making predictions
for points in the remaining element. As a result, each observed data
point is used in the construction of $k-1$ different approximations
and is approximated exactly once. The $k$-fold cross validation method
is data-efficient and provides an unbiased estimate of the expected
prediction error \cite{kohavi1995study}, however it should be noted
that neither this method nor others can provide a universally unbiased
estimator for the variance of prediction error \cite{bengio2004no}.

In addition to the figures displaying approximation results for each
data set, tables of accompanying numerical results are located in the
Appendix. All of the test data sets capture underlying functions that
are almost certainly stochastic. As described in Section
\ref{sec:introduction}, regression techniques appear most appropriate
for these problems. However, in high dimension data grows
exponentially more sparse. Given sparse data, regressors tend towards
interpolation.

%% -------------------------------------------------------------------
\subsection{Forest Fire ($n = 504, d = 12$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-forest-fire.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Forest First Area Burned};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of forest fire area burned under
    recorded weather conditions. The data is presented on a natural
    log scale because most values are small with exponentially fewer
    fires on record that burn large areas.}
  \label{fig:hist-forest-fire}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-forest-fire.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Area Burned Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of area
    that would be burned given environment conditions. $10$-fold
    cross validation as described in the beginning of Section
    \ref{sec:data} is used to evaluated each algorithm. This results
    in exactly one prediction from each algorithm for each data
    point. These boxes depict the median (middle bar), median $95\%$
    confidence interval (cones), quartiles (box edges), fences at
    $3/2$ interquartile range (whiskers), and outliers (dots) of
    absolute prediction error for each model. Similar to figure
    \ref{fig:hist-forest-fire}, the errors are presented on a natural
    log scale. The complementing numerical data to this figure is
    provided in Table \ref{table:error-forest-fire} in the Appendix.}
  \label{fig:error-forest-fire}
\end{figure}

The forest fire data set \cite{cortez2007data} describes the area of
Montesinho park burned over months of the year along with
environmental conditions. The twelve dimensions being used to model
burn area are the $x$ and $y$ spatial coordinates of burns in the
park, month of year (mapped to a 2D circle), the FFMC, DMC, DC, and
ISI indices (see source for details), the temperature, relative
humidity, wind speed, and outdoor rain. The original analysis of this
data set demonstrated it to be difficult to model, likely due to the
skew in response values.

As suggested by Figure \ref{fig:error-forest-fire}, the SVR has the
lowest absolute prediction errors for 80\% of the data, with ShepMod
and BoxSpline being the nearest overall competitors. The effectiveness
of SVR on this data suggests the underlying function can be defined by
relatively few parameters, as well as the importance of capturing the
low-burn-area data points.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Parkinson's Telemonitoring ($n = 5875, d = 19$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-parkinsons.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Parkinson's Total UPDRS Score};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of the Parkinson's patient total UPDRS
    clinical scores that will be approximated by each algorithm.}
  \label{fig:hist-parkinsons}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-parkinsons.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Total UPDRS Score Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the total UPDRS score
    given audio features from patients' home life. $10$-fold
    cross validation as described in the beginning of Section
    \ref{sec:data} is used to evaluated each algorithm. This results
    in exactly one prediction from each algorithm for each data
    point. These boxes depict the median (middle bar), median $95\%$
    confidence interval (cones), quartiles (box edges), fences at
    $3/2$ interquartile range (whiskers), and outliers (dots) of
    absolute prediction error for each model. The complementing
    numerical data to this is provided in table
    \ref{table:error-parkinsons} in the Appendix.}
  \label{fig:error-parkinsons}
\end{figure}

The second data set for evaluation \cite{tsanas2010accurate} is
derived from a speech monitoring study with the intent to
automatically estimate Parkinson's disease symptom development in
Parkinson's patients. The function to be approximated is a
time-consuming clinical evaluation measure referred to as the UPDRS
score. The total UPDRS score given by a clinical evaluation is
estimated through 16 real numbers generated from biomedical voice
measures of in-home sound recordings.

Figure \ref{fig:error-parkinsons} shows the ShepMod algorithm has the
lowest minimum, first quartile, and median of absolute errors for this
problem, while providing the best prediction 66\% of the time. The MLP
has the lowest third quartile and provides the best prediction for
32\% of approximations. The dominance of ShepMod may be in part to
regular-interval total UPDRS scores provided by clinicians, favoring a
nearest-neighbor strategy of prediction.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Australian Daily Rainfall Volume ($n = 2609, d = 23$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-weather.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Daily Rainfall in Sydney};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of daily rainfall in Sydney,
    Australia. The data is presented on a natural log scale because
    the frequency of larger amounts of rainfall is significantly
    less. There is a peak in occurrence of the value $0$, which has a
    notable effect on the resulting model performance.}
  \label{fig:hist-weather}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-weather.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Sydney Rainfall Tomorrow Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of
    rainfall expected on the next calendar day given various sources
    of local meteorological data. $10$-fold cross validation as
    described in the beginning of Section \ref{sec:data} is used to
    evaluated each algorithm. This results in exactly one prediction
    from each algorithm for each data point. These boxes depict the
    median (middle bar), median $95\%$ confidence interval (cones),
    quartiles (box edges), fences at $3/2$ interquartile range
    (whiskers), and outliers (dots) of absolute prediction error for
    each model. The errors are presented on a natural log scale,
    mimicking the presentation in Figure \ref{fig:hist-weather}. The
    complementing numerical data to this figure is provided in table
    \ref{table:error-weather} in the Appendix.}
  \label{fig:error-weather}
\end{figure}

The third data regarding the total daily rainfall in Sydney, Australia
\cite{williams2009rattle} provides a slightly higher dimension
challenge for the interpolants and regressors. This data is composed
of many meteorological readings from the region on a day including:
min and max temperatures, sunshine, wind speed directions (converted
to coordinates on circle), wind speeds, and humidities throughout the
day, day of the year (converted to coordinates on circle), and the
model must predict the amount of rainfall tomorrow.

While Figure \ref{fig:hist-weather} makes MARS look far better than
other techniques, it only provides the best prediction for 11\% of
points. The MLP has the lowest absolute error for 56\% of points and
LSHEP is best for 28\%. MARS likely achieves such a low first quartile
due to the high occurrence of the value 0 in the data.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Credit Card Transaction Amount ($n = 5562, d = 28$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-credit-card.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Credit Card Transaction Amount};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of credit card transaction
    amounts. The data is presented on a natural log scale. The data
    contains a notable frequency peak around $\$1$ transactions. Fewer
    large purchases are made, but some large purchases are in excess
    of five orders of magnitude greater than the smallest purchases.}
  \label{fig:hist-credit-card}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-credit-card.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Transaction Amount Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the expected
    transaction amount given transformed (and obfuscated) vendor and
    customer-descriptive features. $10$-fold cross validation as
    described in the beginning of Section \ref{sec:data} is used to
    evaluated each algorithm. This results in exactly one prediction
    from each algorithm for each data point. These boxes depict the
    median (middle bar), median $95\%$ confidence interval (cones),
    quartiles (box edges), fences at $3/2$ interquartile range
    (whiskers), and outliers (dots) of absolute prediction error for
    each model. The absolute errors in transaction amount predictions
    are presented on a natural log scale, just as in figure
    \ref{fig:hist-credit-card}. The complementing numerical data to
    this figure is provided in Table \ref{table:error-credit-card} in
    the Appendix.}
  \label{fig:error-credit-card}
\end{figure}

The fourth test data set, and the final with a real-valued range, is a
collection of credit card transactions
\cite{pozzolo2015calibrating}. The provided data carries no direct
real-world meaning, it is the output of a principle component analysis
on the original hidden source data. This obfuscation is done to
protect the information of the credit card users. This data has the
largest dimension of all considered, at 28. A model over this data
predicts the transaction amount given the vector of principle
component information.

As can be seen in Figure \ref{fig:error-credit-card}, the MLP
outperforms all other algorithms at the first, second, third, and
fourth quartiles. The MLP produces the lowest absolute error
prediction for 80\% of transactions, Delaunay bests the remaining
20\%. It is likely that with less data, Delaunay would be the best
performer.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{High Performance Computing I/O ($n = 3016, d = 4$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-throughput.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Read Throughtput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of the raw throughput values recorded during all
    IOzone tests across all system configurations. The distribution is
    skewed right, with few tests having significantly higher
    throughput than most others. The data is presented on a natural
    log scale, however the resulting left-skew reveals the data is not
    precisely distributed according to an exponential.}
  \label{fig:hist-throughput}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-throughput.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Read Throughtput KS Statistic};
  \end{tikzpicture}
  \caption{Above, the models directly capable of predicting
    distributions are applied to predicting the expected CDF of I/O
    throughput at previously unseen system configurations. These
    boxes depict the median (middle bar), median $95\%$ confidence
    interval (cones), quartiles (box edges), fences at $3/2$
    interquartile range (whiskers), and outliers (dots) of KS
    statistic between the observed distribution at each system
    configuration and the predicted distribution. The above figure is
    \textit{not} log-scaled like Figure \ref{fig:hist-throughput}. The
    complimentary numerical data for this figure is provided in
    Appendix Table \ref{table:error-throughput}.}
  \label{fig:error-throughput}
\end{figure}

The final of five data sets is derived from \cite{cameron2019moana},
which provides four-dimensional distribution data by executing the
IOzone benchmark \cite{iozone} on a computer system and varying the
system's file size, record size, thread count, and CPU frequency. At
each configuration, IOzone samples the I/O file-read throughput (in
bytes per second) 150 times. Empirical distribution function points
are computed from each set of 150 executions, which are interpolated
with a piecewise cubic Hermite interpolating polynomial
\cite{fritsch1980monotone} to approximate the CDF. All interpolation
algorithms with the exception of LSHEP are used to approximate these
CDFs from system configurations.

Delaunay achieves the lowest KS statistic for 62\% of approximations,
while Voronoi is best for the remaining 38\%. Figure
\ref{fig:error-throughput} shows that while Delaunay may have more
best predictions, the behavior of Voronoi may be preferable.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth]{Figures/plot-throughput-ks.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {KS Statistic for Predicted vs. Actual};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count of KS Statistic};
  \end{tikzpicture}
  \caption{Histograms of the prediction error for each interpolant
    that produces predictions as convex combinations of observed
    data. $10$-fold cross validation as described in the beginning of
    Section \ref{sec:data} is used to evaluated each algorithm. This
    results in exactly one prediction from each algorithm for each
    data point. The distributions show the KS statistics for the
    predicted throughput distribution versus the actual throughput
    distribution. The four vertical red lines represent commonly used
    $p$-values \{0.05, 0.01, 0.001, 1.0e-6\} respectively. All
    predictions to the right of a red line represent CDF predictions
    that are significantly different (by respective $p$-value) from
    the actual distribution according to the KS test. The numerical
    counterpart to this figure is presented in table
    \ref{table:null-hypothesis-results}.}
  \label{fig:throughput-prediction}
\end{figure}

\begin{table}
  \renewcommand{\arraystretch}{1.3}
  \centering
  \begin{tabular}{c|c|c|c|c}
     & $p$ = .05 & $p$ = .01 & $p$ = .001 & $p$ = 1.0e-6\\
    \hline
    Delaunay & $\mathbf{50.3}\%$ & $\mathit{43.5}\%$ & $\mathit{36.2}\%$ & $\mathit{24.7}\%$\\
    ShepMod & $\mathit{51.4}\%$ & $44.8\%$ & $38.1\%$ & $27.7\%$\\
    Voronoi & $52.6\%$ & $\mathbf{43.4}\%$ & $\mathbf{34.4}\%$ & $\mathbf{19.1}\%$\\
    BoxMesh & $99.5\%$ & $98.5\%$ & $96.6\%$ & $89.5\%$\\
  \end{tabular}
  \caption{This is the numerical counterpart to the histogram data
    presented in Figure \ref{fig:throughput-prediction}. The columns
    display the percent of null hypothesis rejections by the KS-test
    when provided different selections of $p$-values for each
    algorithm. The algorithm with the lowest rejection rate at each
    $p$ is boldface, while the second lowest is italicized.}
  \label{table:null-hypothesis-results}
\end{table}

%% ===================================================================

Figure \ref{fig:throughput-prediction} expands on the KS statistic
results presented in Figure \ref{fig:error-throughput}. Agglomerate
errors for each algorithm resemble a Gamma distribution. The
percentages of significant prediction errors with varying $p$-values
are on display in Table \ref{table:null-hypothesis-results}. When
considering the $p=0.001$ results for each technique, a little over
one third of the predicted CDFs are significantly different from the
measured (and presumed) correct CDFs. However, it should be noted that
with 150 samples, the error of an empirical distribution function
(EDF) can reasonably be upwards of $.1$, which serves as a rough
estimate for the lower limit of achievable error. Globally, only a
third of Voronoi predictions fail to capture \textit{all} of the
characteristics of the CDFs at new system configurations.

\section{Discussion}
\label{sec:discussion}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c}
    \hline
    Algorithm & Avg. \% Best & Avg. Fit Time & Avg. App. Time (1)\\
    \hline
    MARS & $4.6\%$ & $22.3$s & $1.37 \times 10^{-3}$s\\
    SVR & $\mathit{21.1}\%$ & $0.503$s & $\mathit{1.37 \times 10^{-4}}$s\\
    MLP & $\mathbf{42.1}\%$ & $210.0$s & $1.02 \times 10^{-3}$s\\
    Delaunay & $6.0\%$ & $\mathbf{1.43 \times 10^{-4}}$s & $3.03$s\\
    ShepMod & $18.4\%$ & $0.711$s & $1.49 \times 10^{-4}$s\\
    LSHEP & $8.4\%$ & $1.69$s & $\mathbf{1.16 \times 10^{-4}}$s\\
    Voronoi & $0.5\%$ & $\mathit{0.127}$s & $0.0375$s\\
    BoxSplines & $0.2\%$ & $0.816$s & $5.26 \times 10^{-4}$s\\
    \hline
  \end{tabular}
  \caption{This data is the average of Appendix tables
    \ref{table:error-forest-fire}, \ref{table:error-parkinsons},
    \ref{table:error-weather}, \ref{table:error-credit-card}, and
    provides a gross summary of overall results. The columns display
    (weighted equally by data set, \textit{not} points) the average
    frequency with which any algorithm provided the lowest absolute
    error approximation, the average time-to-fit, and the average time
    required to approximate one point. Interpolants provide the
    lowest error approximation one third of the time across all data,
    while regressors occupy the other two thirds. This result is
    obtained without any customized tuning or preprocessing to
    maximize the performance of any given algorithm. In practice,
    tuning and preprocessing may have large effects on approximation
    performance.}
  \label{table:avg-performance}
\end{table}

Table \ref{table:avg-performance} summarizes results across the four
test data sets with real-valued ranges. The interpolants discussed in
this paper produce the \textit{best} approximations roughly one third
of the time, and produce competitive approximations for almost all
data sets. These test problems are almost certainly
\textit{stochastic} in nature, but the high dimension yields to
greater sparsity and makes interpolation a competitive approximation
technique.

The major advantages to interpolation lie in the near absence of
\textit{fit} time. Delaunay has no preparations required, LSHEP and
ShepMod require pairwise distance calculations to determine the radii
of influence for points in each model. At least hundreds, and
sometimes hundreds of thousands of predictions can be made by the
interpolants before the most widely used regressor (the MLP) finishes
fitting these relatively small data sets. However, the computational
complexities of all interpolants presented are greater than linear in
either dimension or number of points, meaning large enough data (in
dimension and number of observations) would cause interpolation to be
slower.

The theoretical results presented in Section \ref{sec:theory} directly
apply to Delaunay interpolation, however the performance of Delaunay
does not appear significantly better than other algorithms on these
data sets. This observation may be due to the stochastic nature of the
data, but it also speaks to the alignment of the geometry of the
underlying functions with the approximations generated by the
different interpolation methods. The strong performance of other
interpolants suggests that theoretical results similar to those
presented in this work can be achieved for the other interpolants
under reasonable assumptions.

Finally, most of the interpolants presented in this work benefit from
the ability to model any function over a real vector space with a
range that is closed under convex combinations. The results of the
distribution prediction case study indicate that interpolants can
effectively predict CDFs. The error analysis for that work relies on
the KS statistic, which captures the worst part of any prediction and
hence provides a conservatively large estimate of approximation
error. The average absolute errors in the predicted CDFs are always
lower than the KS statistics. However, the KS statistic was chosen
because of the important statistical theory surrounding it as an error
measure. Considering this circumstance, a nonnegligible volume of
predictions provide impressively low levels of error in that final
case study.

\section{Conclusion}
\label{sec:conclusion}

This work presents new theoretical results for interpolation in
arbitrary dimension with simplicial approximations. Coinciding with
the theoretical results, an empirical evaluation across real-world
problems demonstrates that interpolants produce competitively accurate
models of multivariate phenomenon when compared with common regressors
for sparse, moderately high dimensional problems.

The various interpolants discussed in this paper have been
demonstrated to effectively approximate multivariate phenomena in up
to $30$ dimensions. The underlying constructions are theoretically
straightforward, yet powerful and flexible. Most of the interpolants'
computational complexities make them particularly suitable for
applications in even higher dimension. The major benefits of
interpolation are seen when only a small number of approximations
($\leq 1000$) are made from data and when there are relatively few
data points for the dimension (for empirical results presented, $\log_d n \leq 5$).

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

%%====================================================================
%%====================================================================

\bibliographystyle{spmpsci}      % mathematics and physical sciences
\bibliography{paper}

\newpage

\begin{appendix}

\section*{Appendix}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $0.00984$ & $3.11$ & $7.01$ & $\mathit{11.7}$ & $1090.0$\\
    SVR & $\mathit{0.00402}$ & $\mathbf{0.243}$ & $\mathbf{0.416}$ & $\mathbf{6.19}$ & $1090.0$\\
    MLP & $0.0426$ & $2.63$ & $\mathit{5.27}$ & $14.0$ & $1090.0$\\
    Delaunay & $\mathbf{0.00}$ & $1.98$ & $5.37$ & $13.1$ & $\mathit{1080.0}$\\
    ShepMod & $\mathbf{0.00}$ & $\mathit{1.93}$ & $6.27$ & $16.0$ & $1090.0$\\
    LSHEP & $0.0400$ & $2.17$ & $8.87$ & $19.1$ & $\mathbf{1070.0}$\\
    Voronoi & $0.00982$ & $3.65$ & $7.56$ & $15.6$ & $1090.0$\\
    BoxSpline & $\mathbf{0.00}$ & $2.27$ & $5.91$ & $16.4$ & $1090.0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-forest-fire}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    second quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-forest-fire}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit Time & App. Time (1) & Total App. Time\\
    \cline{1-2}\cline{4-6}
    MARS & $\mathit{7.7}\%$ &  & $29.1$s & $1.37 \times 10^{-3}$s & $0.0686$s\\
    SVR & $\mathbf{80.2}\%$ &  & $5.84 \times 10^{-3}$s & $\mathit{6.20 \times 10^{-5}}$s & $\mathit{3.10 \times 10^{-3}}$s\\
    MLP & $0.0\%$ &  & $32.8$s & $8.71 \times 10^{-4}$s & $0.0436$s\\
    Delaunay & $3.5\%$ &  & $\mathbf{2.49 \times 10^{-5}}$s & $0.0385$s & $1.93$s\\
    ShepMod & $3.7\%$ &  & $6.34 \times 10^{-3}$s & $6.44 \times 10^{-5}$s & $3.22 \times 10^{-3}$s\\
    LSHEP & $5.1\%$ &  & $0.0275$s & $\mathbf{4.63 \times 10^{-5}}$s & $\mathbf{2.31 \times 10^{-3}}$s\\
    Voronoi & $0.0\%$ &  & $\mathit{1.82 \times 10^{-4}}$s & $3.96 \times 10^{-4}$s & $0.0198$s\\
    BoxSpline & $0.4\%$ &  & $7.24 \times 10^{-3}$s & $9.78 \times 10^{-5}$s & $4.89 \times 10^{-3}$s\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating forest fire data in table
    \ref{table:error-forest-fire}. On the right columns are median fit
    time of 454 points, median time for one approximation, and median
    time approximating 50 points.}
  \label{table:best-forest-fire}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $0.00948$ & $9.98$ & $20.4$ & $32.9$ & $863.0$\\
    SVR & $0.00138$ & $3.17$ & $7.31$ & $12.6$ & $\mathbf{27.5}$\\
    MLP & $2.39 \times 10^{-5}$ & $\mathit{0.533}$ & $\mathit{1.25}$ & $\mathbf{2.84}$ & $39.3$\\
    Delaunay & $\mathit{3.72 \times 10^{-12}}$ & $1.20$ & $3.50$ & $7.67$ & $30.7$\\
    ShepMod & $\mathbf{0.00}$ & $\mathbf{0.255}$ & $\mathbf{0.908}$ & $\mathit{3.43}$ & $34.5$\\
    LSHEP & $0.00254$ & $2.93$ & $7.16$ & $13.1$ & $\mathit{29.0}$\\
    Voronoi & $\mathbf{0.00}$ & $1.29$ & $3.52$ & $6.87$ & $30.1$\\
    BoxSpline & $0.00287$ & $4.39$ & $9.10$ & $14.8$ & $41.3$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-parkinsons}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    second quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-parkinsons}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit Time & App. Time (1) & Total App. Time\\
    \cline{1-2}\cline{4-6}
    MARS & $0.0\%$ &  & $37.9$s & $2.53 \times 10^{-3}$s & $1.48$s\\
    SVR & $0.1\%$ &  & $0.859$s & $\mathit{1.81 \times 10^{-4}}$s & $\mathit{0.106}$s\\
    MLP & $\mathit{32.0}\%$ &  & $348.0$s & $1.11 \times 10^{-3}$s & $0.653$s\\
    Delaunay & $0.0\%$ &  & $\mathbf{1.90 \times 10^{-4}}$s & $3.69$s & $2.17 \times 10^{3}$s\\
    ShepMod & $\mathbf{66.4}\%$ &  & $1.13$s & $1.82 \times 10^{-4}$s & $0.107$s\\
    LSHEP & $0.0\%$ &  & $2.39$s & $\mathbf{1.44 \times 10^{-4}}$s & $\mathbf{0.0845}$s\\
    Voronoi & $1.6\%$ &  & $\mathit{0.298}$s & $0.0274$s & $16.1$s\\
    BoxSpline & $0.0\%$ &  & $1.26$s & $6.43 \times 10^{-4}$s & $0.377$s\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating parkinsons data in table
    \ref{table:error-parkinsons}. On the right columns are median fit
    time of 5288 points, median time for one approximation, and median
    time approximating 587 points.}
  \label{table:best-parkinsons}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $\mathit{6.45 \times 10^{-15}}$ & $\mathbf{2.70 \times 10^{-14}}$ & $1.66$ & $1.96$ & $\mathit{53.3}$\\
    SVR & $0.0420$ & $0.143$ & $0.148$ & $\mathit{0.860}$ & $119.0$\\
    MLP & $6.89 \times 10^{-5}$ & $0.0337$ & $\mathbf{0.0795}$ & $\mathbf{0.264}$ & $\mathbf{5.31}$\\
    Delaunay & $\mathbf{0.00}$ & $0.187$ & $0.919$ & $2.72$ & $56.3$\\
    ShepMod & $\mathbf{0.00}$ & $0.0957$ & $0.685$ & $2.90$ & $73.2$\\
    LSHEP & $\mathbf{0.00}$ & $\mathit{0.0153}$ & $\mathit{0.106}$ & $1.17$ & $113.0$\\
    Voronoi & $\mathbf{0.00}$ & $0.430$ & $1.28$ & $2.94$ & $83.8$\\
    BoxSpline & $\mathbf{0.00}$ & $0.789$ & $2.06$ & $5.25$ & $119.0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-weather}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    second quartile, and maximum absolute errors respectively. The
    minimum value of each column is boldface, while the second lowest
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-weather}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit Time & App. Time (1) & Total App. Time\\
    \cline{1-2}\cline{4-6}
    MARS & $10.7\%$ &  & $0.151$s & $1.17 \times 10^{-4}$s & $0.0304$s\\
    SVR & $4.1\%$ &  & $0.133$s & $\mathbf{9.47 \times 10^{-5}}$s & $\mathbf{0.0246}$s\\
    MLP & $\mathbf{56.8}\%$ &  & $169.0$s & $1.37 \times 10^{-3}$s & $0.356$s\\
    Delaunay & $0.1\%$ &  & $\mathbf{1.16 \times 10^{-4}}$s & $1.55$s & $403.0$s\\
    ShepMod & $3.5\%$ &  & $0.265$s & $1.28 \times 10^{-4}$s & $0.0333$s\\
    LSHEP & $\mathit{28.4}\%$ &  & $0.874$s & $\mathit{9.75 \times 10^{-5}}$s & $\mathit{0.0254}$s\\
    Voronoi & $0.2\%$ &  & $\mathit{0.0112}$s & $0.0270$s & $7.01$s\\
    BoxSpline & $0.3\%$ &  & $0.330$s & $4.06 \times 10^{-4}$s & $0.106$s\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating Sydney rainfall data in table
    \ref{table:error-weather}. On the right columns are median fit
    time of 2349 points, median time for one approximation, and median
    time approximating 260 points.}
  \label{table:best-weather}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    MARS & $5.36$ & $3610.0$ & $4580.0$ & $5450.0$ & $13400.0$\\
    SVR & $0.0138$ & $8.47$ & $15.4$ & $41.9$ & $7700.0$\\
    MLP & $0.00151$ & $\mathbf{1.60}$ & $\mathbf{3.86}$ & $\mathbf{8.34}$ & $\mathbf{604.0}$\\
    Delaunay & $\mathbf{0.00}$ & $\mathit{2.69}$ & $\mathit{13.8}$ & $\mathit{35.0}$ & $\mathit{4840.0}$\\
    ShepMod & $\mathbf{0.00}$ & $4.21$ & $17.3$ & $51.6$ & $6510.0$\\
    LSHEP & $1.27$ & $199.0$ & $260.0$ & $343.0$ & $6530.0$\\
    Voronoi & $\mathit{2.89 \times 10^{-10}}$ & $14.7$ & $32.8$ & $52.9$ & $4860.0$\\
    BoxSpline & $0.00740$ & $20.8$ & $40.9$ & $79.6$ & $7640.0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-credit-card}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    second quartile, and maximum absolute errors respectively. The
    minimum value of each column is boldface, while the second lowest
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-credit-card}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit Time & App. Time (1) & Total App. Time\\
    \cline{1-2}\cline{4-6}
    MARS & $0.0\%$ &  & $22.0$s & $1.48 \times 10^{-3}$s & $0.820$s\\
    SVR & $0.0\%$ &  & $1.01$s & $\mathit{2.10 \times 10^{-4}}$s & $\mathit{0.117}$s\\
    MLP & $\mathbf{79.5}\%$ &  & $290.0$s & $7.14 \times 10^{-4}$s & $0.397$s\\
    Delaunay & $\mathit{20.5}\%$ &  & $\mathbf{2.43 \times 10^{-4}}$s & $6.83$s & $3.80 \times 10^{3}$s\\
    ShepMod & $0.1\%$ &  & $1.45$s & $2.20 \times 10^{-4}$s & $0.122$s\\
    LSHEP & $0.0\%$ &  & $3.47$s & $\mathbf{1.76 \times 10^{-4}}$s & $\mathbf{0.0981}$s\\
    Voronoi & $0.0\%$ &  & $\mathit{0.197}$s & $0.0950$s & $52.8$s\\
    BoxSpline & $0.0\%$ &  & $1.66$s & $9.56 \times 10^{-4}$s & $0.532$s\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating credit card transaction data
    in Table \ref{table:error-credit-card}. On the right columns are
    median fit time of 5006 points, median time for one approximation,
    and median time approximating 556 points.}
  \label{table:best-credit-card}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    Algorithm & Min & $25^{th}$ & $50^{th}$ & $75^{th}$ & Max\\
    \hline
    Delaunay & $\mathbf{0.0287}$ & $\mathbf{0.0914}$ & $\mathbf{0.158}$ & $\mathit{0.308}$ & $\mathit{0.897}$\\
    ShepMod & $0.0307$ & $\mathit{0.0983}$ & $\mathit{0.164}$ & $0.335$ & $1.00$\\
    Voronoi & $\mathit{0.0303}$ & $0.100$ & $0.165$ & $\mathbf{0.274}$ & $\mathbf{0.762}$\\
    BoxSpline & $0.0703$ & $0.417$ & $0.561$ & $0.745$ & $1.00$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-throughput}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    second quartile, and maximum KS statistics respectively between
    truth and guess for models predicting the distribution of I/O
    throughput that will be observed at previously unseen system
    configurations. The minimum value of each column is boldface,
    while the second lowest is italicized. All values are rounded to
    three significant digits.}
  \label{table:error-throughput}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c| c |c|c|c|}
    \cline{1-2}\cline{4-6}
    Algorithm & \% Best &  & Fit Time & App. Time (1) & Total App. Time\\
    \cline{1-2}\cline{4-6}
    Delaunay & $\mathbf{62.0}\%$ &  & $\mathbf{5.17 \times 10^{-4}}$s & $0.353$s & $106.0$s\\
    ShepMod & $0.0\%$ &  & $0.0884$s & $\mathbf{1.45 \times 10^{-4}}$s & $\mathbf{0.0436}$s\\
    Voronoi & $\mathit{38.0}\%$ &  & $\mathit{0.0173}$s & $2.53 \times 10^{-3}$s & $0.762$s\\
    BoxSpline & $0.0\%$ &  & $0.0972$s & $\mathit{2.10 \times 10^{-4}}$s & $\mathit{0.0633}$s\\
    \cline{1-2}\cline{4-6}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest KS statistic on the I/O throughput distribution data in
    Table \ref{table:error-throughput}. On the right columns are
    median fit time of 2715 points, median time for one approximation,
    and median time approximating 301 points.}
  \label{table:best-throughput}
\end{table}

\end{appendix}

\end{document}