### 2019-05-03 Disseration/prelim.tex

\documentclass[doublespace,nopageskip]{VTthesis} % nopageskip = Removes arbitrary blank pages.

%% HPC
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{mathptmx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{float}    % For controlling the placement of tables.
%% 
\usepackage{booktabs} % For formal tables
\usepackage{tikz}     % For drawing example boxes
\usetikzlibrary{positioning}

\usepackage{hyphenat}
\hyphenation{op-tical net-works semi-conduc-tor}

%% Command for creating multiplie-line cells inside a table environment
\newcommand{\multilinecell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

% Using the following header instead will create a draft copy of your thesis
%\documentclass[doublespace,draft]{VTthesis}

% Title of your thesis
\title{Interpolation for Modeling and Predicting Variability}

% You should include 3-5 keywords, separated by commas
\keywords{Approximation Theory, Numerical Analysis, High Performance Computing, Computer Security, Nonparametric Statistics, Mathematical Software}

% Your name, including middle initial(s)
\author{Thomas C. H. Lux}

% Change this to your program, e.g. Physics, Civil Engineering, etc.
\program{Computer Science}

% Change this to your degree, e.g. Master of Science, Master of Art, etc.
\degree{Doctor of Philosophy}

% This should be your defense date:
\submitdate{May 16, 2019}

% Committee members. Only have five readers and one chair available.
% Only use the ones you need and don't include the ones you don't need.
% You can also declare a Co-advisor. If you do, the principal and co-advisors
% will be listed as co-advisors on the title page.  Per the VT ETD standards, 
% you should not include titles or educational qualifications such as PhD or Dr.
% You should, however, include middle initials if possible.
\principaladvisor{Layne Watson}
\coadvisor{Yili Hong}
\firstreader{Kirk Cameron}
\secondreader{Bert Huang}
\thirdreader{Gang Wang}
\fourthreader{Yang Cao}

% The dedication and acknowledgement pages are optional. Comment them out to remove them.
%% \dedication{...}
%% \acknowledge{...}

% The abstract is required and should be <=250 words for thesis, <=350 words for dissertation.
\abstract{Approximation is an important problem. This work presents applications of interpolants to modeling random variables. Specifically, this work studies the prediction of distributions of random variables applied to computer system throughput variability. New methods for interpolating data with many dimensions are presented and novel theoretical error bounds are constructed for piecewise linear interpolation. Finally, a mathematical software for approximating distributions from data samples via monotone quintic splines is proposed.}

% The general audience abstract is required. There are currently no word limits.
\abstractgenaud{It is common for scientists to collect data on something they are studying. Often scientists want to create a (predictive) model of that phenomenon based on the data, but the choice of \textit{how} to model the data is a difficult one to answer. This work proposes methods for modeling data that operate under very few assumptions that are broadly applicable across science. Finally, a software package is proposed that would allow scientists to better understand the \textit{true} distribution of their data given relatively few observations.}

\begin{document}
% The following lines set up the front matter of your thesis or dissertation and is required to ensure proper formatting per the VT ETD standards. 
\frontmatter
\maketitle
\tableofcontents

% The list of figures and tables are now optional per the official ETD standards.  Unless you have a very good reason for removing them, you should leave these lists in the document. Comment them out to remove them.
\listoffigures
\listoftables

%% \printnomenclature %creates a list of abbreviations. Comment out to remove it. 
%% ABBR is described here.
%% \nomenclature{ABBR}{abbreviation}

% The following sets up the document for the main part of the thesis or dissertation. Do not comment out or remove this line.
\mainmatter

%% Introduction of variability modeling, the importance of modeling in science in general, then dive into the specific application of modeling techniques to systems (and security).
\chapter{The Importance and Applications of Variability} \label{ch:apps}
\input{apps}

%% Describe all of the interpolants and regressors that will be used throughout this work. (except meshes)
\chapter{Algorithms for Constructing Approximations} \label{ch:algs}
\input{algs}

%% Demonstrate the naive application of interpolants and regressors to data. Discuss results.
\chapter{Na\"{\i}ve Approximations of Variability} \label{ch:naive}
\input{naive}

%% Introduce the theory on box splines. Introduce the algorithms that rely on box splines. Show the results of the code that produces box splines.
%% Demonstrate the application of these meshes (and box splines particularly) to approximation problems. Discuss results.
\chapter{Box Splines: Uses, Constructions, and Applications} \label{ch:boxes}
\input{boxes}

%% Introduce a measurement for error when predicting distribution outputs, then show what happens when system behavior is predicted as a distribution.
\chapter{Stronger Approximations of Variability} \label{ch:strong}
\input{strong}

%% Construct error theory that works for all observed applications to this point.
\chapter{An Error Bound on Piecewise Linear Interpolation} \label{ch:error}
%% Apply approximation algorithms to a final set of problems. Discuss results.
\input{error}

%% Propose new method to improve predictions for distributions.
\chapter{Improving Variability Estimates with Monotone $C^2$ Splines} \label{ch:splines}
\input{splines}

%% Bibliography.
\bibliography{prelim}
\bibliographystyle{plainnat}

% This formats the chapter name to appendix to properly define the headers:
\appendix

% Add your appendices here. You must leave the appendices enclosed in the appendices environment in order for the table of contents to be correct.
\begin{appendices}
  \chapter{Error Bound Appendices} \label{app:error}

  The tables that follow show the precise experimental results for all
  data sets presented in Section \ref{sec:data}. The tests were all run
  serially on an otherwise idle machine with a CentOS 6.10 operating
  system and an Intel i7-3770 CPU operating at 3.4 GHz. The detailed
  performance results in the tables that follow are very much dependent
  on the problem and the algorithm implementation (e.g., some codes are
  TOMS software, some industry distributions, and others are from
  conference paper venues). Different typeface is used to show best
  performers, however not much significance should be attached to
  ranking algorithms based on small time (millisecond) differences. The
  results serve as a demonstration of conceptual validity.

  \begin{table}[H]
    \centering
    \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
      \hline
      Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
      \hline
      MARS & $0$&$00984$ & $3$&$11$ & $7$&$01$ & $\mathit{11}$&$\mathit{7}$ & $1090$&$0$\\
      SVR & $\mathit{0}$&$\mathit{00402}$ & $\mathbf{0}$&$\mathbf{243}$ & $\mathbf{0}$&$\mathbf{416}$ & $\mathbf{6}$&$\mathbf{19}$ & $1090$&$0$\\
      MLP & $0$&$0426$ & $2$&$63$ & $\mathit{5}$&$\mathit{27}$ & $14$&$0$ & $1090$&$0$\\
      Delaunay & $\mathbf{0}$&$\mathbf{00}$ & $1$&$98$ & $5$&$37$ & $13$&$1$ & $\mathit{1080}$&$\mathit{0}$\\
      ShepMod & $\mathbf{0}$&$\mathbf{00}$ & $\mathit{1}$&$\mathit{93}$ & $6$&$27$ & $16$&$0$ & $1090$&$0$\\
      LSHEP & $0$&$0400$ & $2$&$17$ & $8$&$87$ & $19$&$1$ & $\mathbf{1070}$&$\mathbf{0}$\\
      Voronoi & $0$&$00982$ & $3$&$65$ & $7$&$56$ & $15$&$6$ & $1090$&$0$\\
      BoxSpline & $\mathbf{0}$&$\mathbf{00}$ & $2$&$27$ & $5$&$91$ & $16$&$4$ & $1090$&$0$\\
      \hline
    \end{tabular}
    \caption{This numerical data accompanies the visual provided in
      Figure \ref{fig:error-forest-fire}. The columns of absolute error
      percentiles correspond to the minimum, first quartile, median,
      third quartile, and maximum absolute errors respectively. The
      minimum of each column is boldface, while the second lowest value
      is italicized. All values are rounded to three significant
      digits.}
    \label{table:error-forest-fire}
  \end{table}

  \begin{table}[H]
    \centering
    \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
      \cline{1-3}\cline{5-10}
      Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
      \cline{1-3}\cline{5-10}
      MARS & \quad$\mathit{7}$&$\mathit{7}$ &  & \quad\quad$29$&$1$ & \quad$0$&$00137$ & \quad\quad\quad$0$&$0686$\\
      SVR & $\mathbf{80}$&$\mathbf{2}$ &  & $\mathbf{0}$&$\mathbf{00584}$ & $\mathit{0}$&$\mathit{0000620}$ & $\mathit{0}$&$\mathit{00310}$\\
      MLP & $0$&$0$ &  & $32$&$8$ & $0$&$000871$ & $0$&$0436$\\
      Delaunay & $3$&$5$ &  & $0$&$0151$ & $0$&$0234$ & $1$&$18$\\
      ShepMod & $3$&$7$ &  & $\mathit{0}$&$\mathit{00634}$ & $0$&$0000644$ & $0$&$00322$\\
      LSHEP & $5$&$1$ &  & $0$&$0275$ & $\mathbf{0}$&$\mathbf{0000463}$ & $\mathbf{0}$&$\mathbf{00231}$\\
      Voronoi & $0$&$0$ &  & $0$&$0152$ & $0$&$000396$ & $0$&$0198$\\
      BoxSpline & $0$&$4$ &  & $0$&$00724$ & $0$&$0000978$ & $0$&$00489$\\
      \cline{1-3}\cline{5-10}
    \end{tabular}
    \caption{The left above shows how often each algorithm had the
      lowest absolute error approximating forest fire data in Table
      \ref{table:error-forest-fire}. On the right columns are median fit
      time of 454 points, median time for one approximation, and median
      time approximating 50 points.}
    \label{table:best-forest-fire}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
      \hline
      Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
      \hline
      MARS & $0$&$00948$ & $9$&$98$ & $20$&$4$ & $32$&$9$ & $863$&$0$\\
      SVR & $0$&$00138$ & $3$&$17$ & $7$&$31$ & $12$&$6$ & $\mathbf{27}$&$\mathbf{5}$\\
      MLP & $0$&$0000239$ & $\mathit{0}$&$\mathit{533}$ & $\mathit{1}$&$\mathit{25}$ & $\mathbf{2}$&$\mathbf{84}$ & $39$&$3$\\
      Delaunay & $\mathit{3}$&$\mathit{72 \times 10^{-12}}$ & $1$&$20$ & $3$&$50$ & $7$&$67$ & $30$&$7$\\
      ShepMod & $\mathbf{0}$&$\mathbf{0}$ & $\mathbf{0}$&$\mathbf{255}$ & $\mathbf{0}$&$\mathbf{908}$ & $\mathit{3}$&$\mathit{43}$ & $34$&$5$\\
      LSHEP & $0$&$00254$ & $2$&$93$ & $7$&$16$ & $13$&$1$ & $\mathit{29}$&$\mathit{0}$\\
      Voronoi & $\mathbf{0}$&$\mathbf{0}$ & $1$&$29$ & $3$&$52$ & $6$&$87$ & $30$&$1$\\
      BoxSpline & $0$&$00287$ & $4$&$39$ & $9$&$10$ & $14$&$8$ & $41$&$3$\\
      \hline
    \end{tabular}
    \caption{This numerical data accompanies the visual provided in
      Figure \ref{fig:error-parkinsons}. The columns of absolute error
      percentiles correspond to the minimum, first quartile, median,
      third quartile, and maximum absolute errors respectively. The
      minimum of each column is boldface, while the second lowest value
      is italicized. All values are rounded to three significant
      digits.}
    \label{table:error-parkinsons}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
      \cline{1-3}\cline{5-10}
      Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
      \cline{1-3}\cline{5-10}
      MARS & \quad\,$0$&$0$ &  & \quad\quad\quad$37$&$9$ & \quad\,\,$0$&$00253$ & \quad\quad\quad\,\,$1$&$48$\\
      SVR & $0$&$1$ &  & $\mathbf{0}$&$\mathbf{859}$ & $\mathit{0}$&$\mathit{000181}$ & $\mathit{0}$&$\mathit{106}$\\
      MLP & $\mathit{32}$&$\mathit{0}$ &  & $348$&$0$ & $0$&$00111$ & $0$&$653$\\
      Delaunay & $0$&$0$ &  & $2$&$47$ & $1$&$22$ & $720$&$0$\\
      ShepMod & $\mathbf{66}$&$\mathbf{4}$ &  & $\mathit{1}$&$\mathit{13}$ & $0$&$000182$ & $0$&$107$\\
      LSHEP & $0$&$0$ &  & $2$&$39$ & $\mathbf{0}$&$\mathbf{000144}$ & $\mathbf{0}$&$\mathbf{0845}$\\
      Voronoi & $1$&$6$ &  & $2$&$77$ & $0$&$0274$ & $16$&$1$\\
      BoxSpline & $0$&$0$ &  & $1$&$26$ & $0$&$000643$ & $0$&$377$\\
      \cline{1-3}\cline{5-10}
    \end{tabular}
    \caption{The left above shows how often each algorithm had the
      lowest absolute error approximating Parkinson's data in Table
      \ref{table:error-parkinsons}. On the right columns are median fit
      time of 5288 points, median time for one approximation, and median
      time approximating 587 points.}
    \label{table:best-parkinsons}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
      \hline
      Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
      \hline
      MARS & $\mathit{6}$&$\mathit{45 \times 10^{-15}}$ & $\mathbf{2}$&$\mathbf{70 \times 10^{-14}}$ & $1$&$66$ & $1$&$96$ & $\mathit{53}$&$\mathit{3}$\\
      SVR & $0$&$0420$ & $0$&$143$ & $0$&$148$ & $\mathit{0}$&$\mathit{860}$ & $119$&$0$\\
      MLP & $0$&$0000689$ & $0$&$0337$ & $\mathbf{0}$&$\mathbf{0795}$ & $\mathbf{0}$&$\mathbf{264}$ & $\mathbf{5}$&$\mathbf{31}$\\
      Delaunay & $\mathbf{0}$&$\mathbf{0}$ & $0$&$187$ & $0$&$919$ & $2$&$72$ & $56$&$3$\\
      ShepMod & $\mathbf{0}$&$\mathbf{0}$ & $0$&$0957$ & $0$&$685$ & $2$&$90$ & $73$&$2$\\
      LSHEP & $\mathbf{0}$&$\mathbf{0}$ & $\mathit{0}$&$\mathit{0153}$ & $\mathit{0}$&$\mathit{106}$ & $1$&$17$ & $113$&$0$\\
      Voronoi & $\mathbf{0}$&$\mathbf{0}$ & $0$&$430$ & $1$&$28$ & $2$&$94$ & $83$&$8$\\
      BoxSpline & $\mathbf{0}$&$\mathbf{0}$ & $0$&$789$ & $2$&$06$ & $5$&$25$ & $119$&$0$\\
      \hline
    \end{tabular}
    \caption{This numerical data accompanies the visual provided in
      Figure \ref{fig:error-weather}. The columns of absolute error
      percentiles correspond to the minimum, first quartile, median,
      third quartile, and maximum absolute errors respectively. The
      minimum value of each column is boldface, while the second lowest
      is italicized. All values are rounded to three significant
      digits.}
    \label{table:error-weather}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
      \cline{1-3}\cline{5-10}
      Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
      \cline{1-3}\cline{5-10}
      MARS & \,\,\,\,$10$&$7$ &  & \quad\quad\quad$\mathit{0}$&$\mathit{151}$ & \quad$0$&$000117$ & \quad\quad\quad\,\,$0$&$0304$\\
      SVR & $4$&$1$ &  & $\mathbf{0}$&$\mathbf{133}$ & $\mathbf{0}$&$\mathbf{0000947}$ & $\mathbf{0}$&$\mathbf{0246}$\\
      MLP & $\mathbf{56}$&$\mathbf{8}$ &  & $169$&$0$ & $0$&$00137$ & $0$&$356$\\
      Delaunay & $0$&$1$ &  & $0$&$664$ & $0$&$886$ & $230$&$0$\\
      ShepMod & $3$&$5$ &  & $0$&$265$ & $0$&$000128$ & $0$&$0333$\\
      LSHEP & $\mathit{28}$&$\mathit{4}$ &  & $0$&$874$ & $\mathit{0}$&$\mathit{0000975}$ & $\mathit{0}$&$\mathit{0254}$\\
      Voronoi & $0$&$2$ &  & $0$&$675$ & $0$&$0270$ & $7$&$01$\\
      BoxSpline & $0$&$3$ &  & $0$&$330$ & $0$&$000406$ & $0$&$106$\\
      \cline{1-3}\cline{5-10}
    \end{tabular}
    \caption{Left table shows how often each algorithm had the lowest
      absolute error approximating Sydney rainfall data in Table
      \ref{table:error-weather}. On the right columns are median fit
      time of 2349 points, median time for one approximation, and median
      time approximating 260 points.}
    \label{table:best-weather}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
      \hline
      Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
      \hline
      MARS & $5$&$36$ & $3610$&$0$ & $4580$&$0$ & $5450$&$0$ & $13400$&$0$\\
      SVR & $0$&$0138$ & $8$&$47$ & $15$&$4$ & $41$&$9$ & $7700$&$0$\\
      MLP & $0$&$00151$ & $\mathbf{1}$&$\mathbf{60}$ & $\mathbf{3}$&$\mathbf{86}$ & $\mathbf{8}$&$\mathbf{34}$ & $\mathbf{604}$&$\mathbf{0}$\\
      Delaunay & $\mathbf{0}$&$\mathbf{0}$ & $\mathit{2}$&$\mathit{69}$ & $\mathit{13}$&$\mathit{8}$ & $\mathit{35}$&$\mathit{0}$ & $\mathit{4840}$&$\mathit{0}$\\
      ShepMod & $\mathbf{0}$&$\mathbf{0}$ & $4$&$21$ & $17$&$3$ & $51$&$6$ & $6510$&$0$\\
      LSHEP & $1$&$27$ & $199$&$0$ & $260$&$0$ & $343$&$0$ & $6530$&$0$\\
      Voronoi & $\mathit{2}$&$\mathit{89 \times 10^{-10}}$ & $14$&$7$ & $32$&$8$ & $52$&$9$ & $4860$&$0$\\
      BoxSpline & $0$&$00740$ & $20$&$8$ & $40$&$9$ & $79$&$6$ & $7640$&$0$\\
      \hline
    \end{tabular}
    \caption{This numerical data accompanies the visual provided in
      Figure \ref{fig:error-credit-card}. The columns of absolute error
      percentiles correspond to the minimum, first quartile, median,
      third quartile, and maximum absolute errors respectively. The
      minimum value of each column is boldface, while the second lowest
      is italicized. All values are rounded to three significant
      digits.}
    \label{table:error-credit-card}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
      \cline{1-3}\cline{5-10}
      Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
      \cline{1-3}\cline{5-10}
      MARS & \quad\,$0$&$0$ &  & \quad\quad\quad$22$&$0$ & \quad$0$&$00148$ & \quad\quad\quad\,\,$0$&$820$\\
      SVR & $0$&$0$ &  & $\mathbf{1}$&$\mathbf{01}$ & $\mathit{0}$&$\mathit{000210}$ & $\mathit{0}$&$\mathit{117}$\\
      MLP & $\mathbf{79}$&$\mathbf{5}$ &  & $290$&$0$ & $0$&$000714$ & $0$&$397$\\
      Delaunay & $\mathit{20}$&$\mathit{5}$ &  & $3$&$12$ & $3$&$71$ & $2070$&$0$\\
      ShepMod & $0$&$1$ &  & $\mathit{1}$&$\mathit{4}5$ & $0$&$000220$ & $0$&$122$\\
      LSHEP & $0$&$0$ &  & $3$&$47$ & $\mathbf{0}$&$\mathbf{000176}$ & $\mathbf{0}$&$\mathbf{0981}$\\
      Voronoi & $0$&$0$ &  & $3$&$32$ & $0$&$0950$ & $52$&$8$\\
      BoxSpline & $0$&$0$ &  & $1$&$66$ & $0$&$000956$ & $0$&$532$\\
      \cline{1-3}\cline{5-10}
    \end{tabular}
    \caption{The left above shows how often each algorithm had the
      lowest absolute error approximating credit card transaction data
      in Table \ref{table:error-credit-card}. On the right columns are
      median fit time of 5006 points, median time for one approximation,
      and median time approximating 556 points.}
    \label{table:best-credit-card}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
      \hline
      Algorithm & \multicolumn{2}{c|}{Min} & \multicolumn{2}{c|}{$25^{th}$} & \multicolumn{2}{c|}{$50^{th}$} & \multicolumn{2}{c|}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
      \hline
      Delaunay & $\mathbf{0}$&$\mathbf{0287}$ & $\mathbf{0}$&$\mathbf{0914}$ & $\mathbf{0}$&$\mathbf{158}$ & $\mathit{0}$&$\mathit{308}$ & $\mathit{0}$&$\mathit{897}$\\
      ShepMod & $0$&$0307$ & $\mathit{0}$&$\mathit{0983}$ & $\mathit{0}$&$\mathit{164}$ & $0$&$335$ & $1$&$00$\\
      Voronoi & $\mathit{0}$&$\mathit{0303}$ & $0$&$100$ & $0$&$165$ & $\mathbf{0}$&$\mathbf{274}$ & $\mathbf{0}$&$\mathbf{762}$\\
      BoxSpline & $0$&$0703$ & $0$&$417$ & $0$&$561$ & $0$&$745$ & $1$&$00$\\
      \hline
    \end{tabular}
    \caption{This numerical data accompanies the visual provided in
      Figure \ref{fig:error-throughput}. The columns of absolute error
      percentiles correspond to the minimum, first quartile, median,
      third quartile, and maximum KS statistics respectively between
      truth and guess for models predicting the distribution of I/O
      throughput that will be observed at previously unseen system
      configurations. The minimum value of each column is boldface,
      while the second lowest is italicized. All values are rounded to
      three significant digits.}
    \label{table:error-throughput}
  \end{table}

  \begin{table}
    \centering
    \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
      \cline{1-3}\cline{5-10}
      Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
      \cline{1-3}\cline{5-10}
      Delaunay & \,\,$\mathbf{62}$&$\mathbf{0}$ &  & \quad\quad\quad$0$&$344$ & \quad$0$&$00984$ & \quad\quad\,\,$2$&$71$\\
      ShepMod & $0$&$0$ &  & $\mathbf{0}$&$\mathbf{0884}$ & $\mathbf{0}$&$\mathbf{000145}$ & $\mathbf{0}$&$\mathbf{0436}$\\
      Voronoi & $\mathit{38}$&$\mathit{0}$ &  & $0$&$360$ & $0$&$00253$ & $0$&$762$\\
      BoxSpline & $0$&$0$ &  & $\mathit{0}$&$\mathit{0972}$ & $\mathit{0}$&$\mathit{000210}$ & $\mathit{0}$&$\mathit{0633}$\\
      \cline{1-3}\cline{5-10}
    \end{tabular}
    \caption{The left above shows how often each algorithm had the
      lowest KS statistic on the I/O throughput distribution data in
      Table \ref{table:error-throughput}. On the right columns are
      median fit time of 2715 points, median time for one approximation,
      and median time approximating 301 points.}
    \label{table:best-throughput}
  \end{table}

\end{appendices}

\end{document}

%****************************************************************************
% Below are some general suggestions for writing your dissertation:
%
% 1. Label everything with a meaningful prefix so that you
%    can refer back to sections, tables, figures, equations, etc.
%    Usage \label{<prefix>:<label_name>} where some suggested
%    prefixes are:
%			ch: Chapter
%     		se: Section
%     		ss: Subsection
%     		sss: Sub-subsection
%			app: Appendix
%     		ase: Appendix section
%     		tab: Tables
%     		fig: Figures
%     		sfig: Sub-figures
%     		eq: Equations
%
% 2. The VTthesis class provides for natbib citations. You should upload
%	 one or more *.bib bibtex files. Suppose you have two bib files: some_refs.bib and 
%    other_refs.bib.  Then your bibliography line to include them
%    will be:
%      \bibliography{some_refs, other_refs}
%    where multiple files are separated by commas. In the body of 
%    your work, you can cite your references using natbib citations.
%    Examples:
%      Citation                     Output
%      -------------------------------------------------------
%      \cite{doe_title_2016}        [18]
%      \citet{doe_title_2016}       Doe et al. [18]
%      \citet*{doe_title_2016}      Doe, Jones, and Smith [18]
%
%    For a complete list of options, see
%      https://www.ctan.org/pkg/natbib?lang=en
%
% 3. Here is a sample table. Notice that the caption is centered at the top. Also
%    notice that we use booktabs formatting. You should not use vertical lines
%    in your tables.
% 
%				\begin{table}[htb]
%					\centering
%					\caption{Approximate computation times in hh:mm:ss for full order 						versus reduced order models.}
%					\begin{tabular}{ccc}
%						\toprule
%						& \multicolumn{2}{c}{Computation Time}\\
%						\cmidrule(r){2-3}
%						$\overline{U}_{in}$ m/s & Full Model & ROM \\
%						\midrule
%						0.90 & 2:00:00 & 2:08:00\\
%						0.88 & 2:00:00 & 0:00:03\\
%						0.92 & 2:00:00 & 0:00:03\\
%						\midrule
%						Total & 6:00:00 & 2:08:06\\
%						\bottomrule
%					\end{tabular}
%					\label{tab:time_rom}
%				\end{table}
% 
% 4. Below are some sample figures. Notice the caption is centered below the
%    figure.
%    a. Single centered figure:
%					\begin{figure}[htb]
%						\centering
%						\includegraphics[scale=0.5]{my_figure.eps}
%						\caption{Average outlet velocity magnitude given an average  
%				        input velocity magnitude of 0.88 m/s.} 
%						\label{fig:output_rom}
%					\end{figure}
%    b. Two by two grid of figures with subcaptions
%					\begin{figure}[htb]
%						\centering
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_1_1.eps}
%							\caption{Subcaption number one}
%							\label{sfig:first_subfig}
%						\end{subfigure}
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_1_2.png}
%							\caption{Subcaption number two}
%							\label{sfig:second_subfig}
%						\end{subfigure}
%
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_2_1.pdf}
%							\caption{Subcaption number three}
%							\label{sfig:third_subfig}
%						\end{subfigure}
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_2_2.eps}
%							\caption{Subcaption number four}
%							\label{sfig:fourth_subfig}
%						\end{subfigure}
%						\caption{Here is my main caption describing the relationship 							between the 4 subimages}
%						\label{fig:main_figure}
%					\end{figure}
%
%----------------------------------------------------------------------------
%
% The following is a list of definitions and packages provided by VTthesis:
%
% A. The following packages are provided by the VTthesis class:
%      amsmath, amsthm, amssymb, enumerate, natbib, hyperref, graphicx, 
%      tikz (with shapes and arrows libraries), caption, subcaption,
%      listings, verbatim
%
% B. The following theorem environments are defined by VTthesis:
%      theorem, proposition, lemma, corollary, conjecture
% 
% C. The following definition environments are defined by VTthesis:
%      definition, example, remark, algorithm
%
%----------------------------------------------------------------------------
%
%  I hope this template file and the VTthesis class will keep you from having 
%  to worry about the formatting and allow you to focus on the actual writing.
%  Good luck, and happy writing.
%    Alan Lattimer, VT, 2016
%
%****************************************************************************