### 2020-06-23 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/1-introduction.tex
Updated code package sent off last night. New tex revisions from Dr. Watson received, implementing.

\bigskip\hrule\bigskip\medskip
\heading{1. INTRODUCTION}

Many domains of science rely on smooth approximations to real-valued
functions over a closed interval. Piecewise polynomial functions
(splines) provide the smooth approximations for animation in graphics
[Herman et al. 2015; Quint 2003], aesthetic structural support in
architecture [Brennan 2020], efficient aerodynamic surfaces in
automotive and aerospace engineering [Brennan 2020], prolonged
effective operation of electric motors [Berglund et al 2009], and
accurate nonparametric approximations in statistics [Knott
  2012]. While polynomial interpolants and regressors apply broadly,
splines are often a good choice because they can approximate globally
complex functions while minimizing the local complexity of an
approximation.

It is often the case that the true underlying function or phenomenon
being modeled has known properties e.g., convexity, positivity,
various levels of continuity, or monotonicity. Given a reasonable
amount of data, it quickly becomes difficult to achieve desirable
properties in a single polynomial function. In general, the
maintenance of function properties through interpolation/regression is
referred to as {\it shape preserving} [Fritsch and Carlson 1980;
  Gregory 1985]. The specific properties the present algorithm will
preserve in approximations are monotonicity and $C^2$ continuity. In
addition to previously mentioned applications, these properties are
crucially important in statistics to the approximation of a cumulative
distribution function and subsequently the effective generation of
random numbers from a specified distribution [Ramsay 1988].  A spline
function with these properties could approximate a cumulative
distribution function to a high level of accuracy with relatively few
intervals. A twice continuously differentiable approximation to a
cumulative distribution function (CDF) would produce a corresponding
probability density function (PDF) that is continuously
differentiable, which is desirable.

The currently available software for monotone piecewise polynomial
interpolation includes quadratic [He and Shi 1998], cubic [Fritsch and
  Carlson 1980], and (with limited application) quartic [Wang and Tan
  2004; Piah and Unsworth 2011; Yao and Nelson 2018] cases. In
addition, a statistical method for bootstrapping the construction of
an arbitrarily smooth monotone fit exists [Leitenstorfer and Tutz
  2006], but the method does not take advantage of known analytic
properties of quintic polynomials. The code by Fritsch [1982] for
$C^1$ cubic spline interpolation is the predominantly utilized code
for constructing monotone interpolants at present. Theory has been
provided for the quintic case [Ulrich and Watson 1994; He{\ss} and
  Schmidt 1994] and that theory was recently utilized in a proposed
algorithm [Lux 2020] for monotone quintic spline construction, however
no published mathematical software exists.

The importance of piecewise quintic interpolation over lower order
approximations can be simply observed. In general, the order of a
polynomial determines the number of function (and derivative) values
it can interpolate, which in turn determines the growth rate of error
away from interpolated values. This work constructs quintic splines in
order to match the function value and two derivatives at either end of
an interval, which requires six degrees of freedom in an interpolating
polynomial. This work provides a Fortran 2003 subroutine {\tt MQSI}
based on the necessary and sufficient conditions in Ulrich and Watson
[1994] for the construction of monotone quintic spline interpolants of
monotone data.

The remainder of this paper is structured as follows: Section 2 provides
the algorithms for constructing a $C^2$ monotone quintic spline interpolant
to monotone data, Section 3 outlines the method of spline representation
($B$-spline basis) and evaluation, Section 4 analyzes the complexity and
sensitivity of the algorithms in {\tt MQSI}, and Section 5 presents an
empirical performance study and some graphs of constructed interpolants.