### 2020-06-23 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/4-analysis.tex

\heading{4. COMPLEXITY AND SENSITIVITY}

Both Algorithms 1 and 2 have ${\cal O}(1)$ runtime. Given a fixed
schedule for shrinking derivative values, Algorithm 3 has a ${\cal
  O}(n)$ runtime for $n$ data points. In execution, the majority of
time is spent solving the banded linear system of equations for
B-spline coefficients. The quadratic facet model produces a unique
sensitivity to input perturbation, as small changes in input may cause
data points to be associated with different local quadratic
interpolants. While this makes the technique susceptible to random
perturbations, it generally produces high quality monotone quintic
splines and is preferred for that reason.