### 2020-06-24 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/3-spline-rep.tex
Reorganized folders, started working on visuals.

\heading{3. SPLINE REPRESENTATION}
The monotone quintic spline interpolant $Q(x)$ is represented in terms of
a B-spline basis. The routine {\tt FIT\_SPLINE} in this package computes
the B-spline coefficients $\alpha_i$ of $Q(x)=\sum_{i=1}^{3n} \alpha_i
B_{i,6,t}(x)$ to match the piecewise quintic polynomial values and (first
two) derivatives at the breakpoints $X_i$, where the spline order is six
and the knot sequence $t$ has the breakpoint multiplicities (6, 3, $\ldots$,
3, 6).  The routine {\tt EVAL\_SPLINE} evaluates a spline represented in
terms of a B-spline basis. A Fortran 2003 implementation {\tt EVAL\_BSPLINE}
of the B-spline recurrence relation evaluation code by C. deBoor [1978]
for the value, derivatives, and integral of a B-spline is also provided.