### 2020-06-26 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/misc/example_tex/old_files/0-conclusions.tex
Changed some directory names and fixed some typos in code files. Still working on the paper.

\heading{5. CONCLUSION AND FUTURE WORK}

This paper proposes and tests an algorithm for constructing monotone
quintic spline interpolants. Experiments demonstrate an improvement in
approximation accuracy over monotone cubic spline interpolants, as
expected based on theory. There are still open avenues of research
going forward, such as an alternative sufficient condition for
enforcing monotonicity or increased order monotone approximations. If
the monotonicity conditions can be generalized to any order or made
linear, the search for a monotone interpolating spline could
potentially be formulated as a convex optimization problem. Finally,
this work could be used to improve cumulative distribution function
(CDF) estimates as well as predictive models that use CDFs.