### 2020-06-26 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/misc/example_tex/old_files/0-discussion.tex

\heading{4. DISCUSSION}

The monotone quintic spline interpolant provides a distinct increase
in accuracy over the monotone cubic variant. The computational
complexity of this algorithm is identical to the existing cubic method
and the cost of evaluating the spline after construction is
unchanged. While the number of computations required per interval has
increased, the number of points required to achieve the same level of
accuracy has decreased. In the present state, the algorithm for
enforcing monotonicity on a spline is not as trivially parallel as the
cubic algorithm, however it can still be parallelized. The checks for
monotonicity across all intervals are independent and the monotonicity
adjustments could be computed independently and intelligently merged
with minor changes to the serial algorithm.

Acknowledging that the algorithm for constructing a monotone quintic
spline interpolant is slightly more expensive than the cubic case,
gains in accuracy or decreased necessary number of points are often
worth the computational effort. In practice the costs of spline
interpolation are often dominated by evaluation rather than
construction, and the increased accuracy is afforded for a negligible
increase in computational cost.