### 2020-07-20 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/4-analysis.tex

\heading{4. COMPLEXITY AND SENSITIVITY}

\topinsert
\centerline{\epsfxsize=4truein \epsffile{vis/1-sensitivity.eps}}
{\everymath={\scriptstyle}
\narrower\noindent\rmVIII Fig.\ 1. A demonstration of the quadratic
  facet model's sensitivity to small data perturbations. This example is
  composed of two quadratic functions $f_1(x) = x^2$ over points $\{1$,
  $2$, $5/2\}$, and $f_2(x) = (x-2)^2 + 6$ over points $\{5/2$, $3$,
  $4\}$. Notably, $f_1(5/2) = f_2(5/2)$ and $f_1$, $f_2$ have the same
  curvature. Given the exact five data points seen above, the quadratic
  facet model produces the slope seen in the solid blue line at $x = 5/2$.
  However, by subtracting the value of $f_3$ $= \epsilon(x-2)^2$ from
  points at $x$ $= \{3$, $4\}$, where $\epsilon$ is the machine precision
  ($2^{-52}$ for an IEEE 64-bit real), the quadratic facet model produces
  the slope seen in the dashed red line at $x = 5/2$. This is the nature
  of a facet model and a side effect of associating data with local facets.
\par} \endinsert

Algorithms 1, 2, and 4 have ${\cal O}(1)$ runtime. Given a fixed
schedule for shrinking derivative values, Algorithm 3 has a ${\cal
  O}(n)$ runtime for $n$ data points. In execution, the majority of
the time, still ${\cal O}(n)$, is spent solving the banded linear
system of equations for the B-spline coefficients. Thus for $n$ data
points, the overall execution time is ${\cal O}(n)$. The quadratic
facet model produces a unique sensitivity to input perturbation, as
small changes in input may cause different quadratic facets to be
associated with a breakpoint, and thus different initial derivative
estimates can be produced. This phenomenon is depicted in Figure
1. Despite this sensitivity, the quadratic facet model is still
preferred because it generally provides aesthetically pleasing (low
{\it wiggle}) initial estimates of the first and second derivatives at
the breakpoints while perfectly capturing local low-order phenomena
(e.g., a run of points on a straight line) in the data. The binary
search for a point near the monotone boundary in $(\tau_1$, $\alpha$,
$\beta$, $\gamma)$ space is preferred because it results in monotone
quintic spline interpolants that are nearer to initial estimates than
a policy that strictly shrinks derivative values.