### 2020-07-23 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/2-monotone-quintic.tex
Made modifications including Dr. Watsons recent revisions. Working towards a submission version.

\heading{2. MONOTONE QUINTIC INTERPOLATION}

In order to construct a monotone quintic interpolating spline, two
primary problems must be solved. First, reasonable derivative values
at data points need to be estimated. Second, the estimated derivative
values need to be modified to enforce monotonicity on all polynomial
pieces.

Fritsch and Carlson [1980] originally proposed the use of central differences
to estimate derivatives, however this often leads to extra and unnecessary
{\it wiggles} in the spline when used to approximate second derivatives.
In an attempt to capture the local shape of the data, this package uses a
facet model from image processing [Haralick and Watson 1981] to estimate
first and second derivatives at breakpoints. Rather than picking a local
linear or quadratic fit with minimal residual, this work uses a quadratic
facet model that selects the local quadratic interpolant with minimum
magnitude curvature.

%% ===================================================================
%%                   Algorithm 1: QUADRATIC_FACET
\vskip 5mm
{\parindent 0mm
{\bf Algorithm 1:} {
\tt QUADRATIC\_FACET}$\bigl(X(1{:}n)$, $Y(1{:}n)$, $i \bigr)$

where $X_j$, $Y_j$ $\in \bbb{R}$ for $j = 1$, $\ldots$, $n$, $1 \le i
\le n$, and $n \ge 3$. Returns the slope and curvature at $X_i$ of the
local quadratic interpolant with minimum magnitude curvature.

}
{\parindent=3mm
%% -------------------------------------------------------------------
\item{} {\tt if} $\bigl((i\ne1\wedge Y_i \approx Y_{i-1})$ {\tt or}
$(i\ne n\wedge Y_i \approx Y_{i+1})\bigr)$ {\tt then return $(0,0)$}

\item{} {\tt else if} $i=1$ {\tt then}
\itemitem{} $f_1\:\hbox{interpolant to }(X_1,Y_1)$,
  $(X_2,Y_2)$, and $(X_3,Y_3)$.
\itemitem{} {\tt if} $\bigl(Df_1(X_1)(Y_2-Y_1)<0\bigr)$ {\tt then return}
  $\bigl(0,0\bigr)$
\itemitem{} {\tt else return} $\bigl(Df_1(X_1),D^2f_1\bigr)$
\itemitem{} {\tt endif}

\item{} {\tt else if} $i=n$ {\tt then}
\itemitem{} $f_1\:\hbox{interpolant to }(X_{n-2},Y_{n-2})$,
  $(X_{n-1},Y_{n-1})$, and $(X_n,Y_n)$.
\itemitem{} {\tt if} $\bigl(Df_1(X_n)(Y_n-Y_{n-1})<0\bigr)$ {\tt then return}
  $\bigl(0,0\bigr)$
\itemitem{} {\tt else return} $\bigl(Df_1(X_n),D^2f_1\bigr)$
\itemitem{} {\tt endif}

\item{} {\tt else if} $\bigl(1<i<n\wedge (Y_{i+1} - Y_i)(Y_i - Y_{i-1}) < 0
\bigr)$ {\tt then}
\itemitem{} The point $(X_i$, $Y_i)$ is an extreme point. The
quadratic with minimum magnitude
curvature that has slope zero at $X_i$ will be the facet chosen.
\itemitem{} $f_1\:\hbox{interpolant to }(X_{i-1},Y_{i-1})$, $(X_i,Y_i)$,
  and $Df_1(X_i) = 0$.
\itemitem{} $f_2\:\hbox{interpolant to }(X_i,Y_i)$, $(X_{i+1},Y_{i+1})$,
  and $Df_2(X_i) = 0$.
\itemitem{} {\tt if} $\bigl(|D^2f_1| \leq |D^2f_2|\bigr)$ {\tt then
  return} $\bigl(0$, $D^2f_1\bigr)$
\itemitem{} {\tt else return} $\bigl(0$, $D^2f_2\bigr)$
\itemitem{} {\tt endif}

\item{} {\tt else}
\itemitem{} {The point $(X_i$, $Y_i)$ is in a monotone segment of
data. In the following, it is possible that $f_1$ or $f_3$ do
not exist because $i \in \{2, n-1\}$. In those cases, the minimum
magnitude curvature among existing quadratics is chosen.}
\itemitem{} $f_1\:\hbox{interpolant to }(X_{i-2},Y_{i-2})$,
  $(X_{i-1},Y_{i-1})$, and $(X_i,Y_i)$.
\itemitem{} $f_2\:\hbox{interpolant to }(X_{i-1},Y_{i-1})$,
  $(X_i,Y_i)$, and $(X_{i+1},Y_{i+1})$.
\itemitem{} $f_3\:\hbox{interpolant to }(X_i,Y_i)$,
  $(X_{i+1},Y_{i+1})$, and $(X_{i+2},Y_{i+2})$.

\itemitem{} {\tt if} $\bigl(Df_1(X_i)(Y_i-Y_{i-1}) \ge 0 \wedge
  |D^2f_1| = \min\bigl\{ |D^2f_1|,
  |D^2f_2|, |D^2f_3|\bigr\} \bigr)$ {\tt then}
\itemitem{} \codent {\tt return} $\bigl(Df_1(X_i),D^2f_1\bigr)$

\itemitem{} {\tt else if} $\bigl(Df_2(X_i)(Y_i-Y_{i-1}) \ge 0 \wedge
  |D^2f_2| = \min\bigl\{ |D^2f_1|,
  |D^2f_2|, |D^2f_3|\bigr\} \bigr)$ {\tt then}
\itemitem{} \codent {\tt return} $\bigl(Df_2(X_i),D^2f_2\bigr)$

\itemitem{} {\tt else if} $\bigl(Df_3(X_i)(Y_{i+1}-Y_i) \ge 0\bigr)$ {\tt then}
\itemitem{} \codent {\tt return} $\bigl(Df_3(X_i),D^2f_3\bigr)$
\itemitem{} {\tt else return} $(0, 0)$
\itemitem{} {\tt endif}
\item{} {\tt endif}
}
\vskip 5mm
%% ----------------------------------------------------------------------

The estimated derivative values by the quadratic facet model are not
guaranteed to produce monotone quintic polynomial segments. Ulrich and
Watson [1994] established tight constraints on the monotonicity of a
quintic polynomial piece, while deferring to He{\ss} and Schmidt
[1994] for a relevant simplified case. The following algorithm
implements a sharp check for monotonicity by considering the
nondecreasing case. The nonincreasing case is handled similarly.

%% ===================================================================
%%                    Algorithm 2: IS_MONOTONE
\vskip 5mm
{\parindent 0mm
{\bf Algorithm 2:}

{\tt IS\_MONOTONE}$\bigl(x_0$, $x_1$, $f(x_0)$, $Df(x_0)$,
$D^2f(x_0)$, $f(x_1)$, $Df(x_1)$, $D^2f(x_1) \bigr)$

where $x_0$, $x_1 \in {\bbb R}$, $x_0 < x_1$, and $f$ is an order six
polynomial defined by $f(x_0)$, $Df(x_0)$, $D^2f(x_0)$, $f(x_1)$,
$Df(x_1)$, $D^2f(x_1)$. Returns {\tt TRUE} if $f$ is monotone
increasing on $[x_0,x_1]$.

}
{\parindent=6mm
%% -------------------------------------------------------------------
\item{1.} {\tt if} $\bigl(f(x_0) \approx f(x_1)\bigr)$ {\tt then}
\item{2.} \codent {\tt return} $\bigl( 0 = Df(x_0) = Df(x_1)
  = D^2f(x_0) = D^2f(x_1) \bigr)$
\item{3.} {\tt endif}
\item{4.} {\tt if} $\bigl(Df(x_0) < 0$ {\tt or} $Df(x_1) < 0\bigr)$ {\tt
  then return FALSE endif}
\item{5.} $w \: x_1 - x_0$
\item{6.} $z \: f(x_1) - f(x_0)$

\item{} {The necessity of Steps 1--4 follows directly from the
  fact that $f$ is $C^2$. The following Steps 7--13 coincide with a
  simplified condition for quintic monotonicity that reduces to one of
  cubic positivity studied by Schmidt and He{\ss} [1988]. Given
  $\alpha$, $\beta$, $\gamma$, and $\delta$ as defined by Schmidt and
  He{\ss}, monotonicity results when $\alpha \geq 0$, $\delta \geq 0$,
  $\beta \geq \alpha - 2 \sqrt{\alpha \delta}$, and $\gamma \geq
  \delta - 2 \sqrt{\alpha \delta}$.  Step 4 checked for $\delta < 0$,
  Step 8 checks $\alpha < 0$, Step 10 checks $\beta < \alpha - 2
  \sqrt{\alpha \delta}$, and Step 11 checks $\gamma < \delta - 2
  \sqrt{\alpha \delta}$. If none of the monotonicity conditions are
  violated, then the order six piece is monotone and Step 12
  concludes.}

\item{7.} {\tt if} $\bigl(Df(x_0) \approx 0$ {\tt or} $Df(x_1) \approx
0\bigr)$ {\tt then}
\item{8.} \codent {\tt if} $\bigl(D^2f(x_1)w > 4Df(x_1)$ {\tt then
return FALSE endif}
\item{9.} \codent $t \: 2 \sqrt{Df(x_0) (4Df(x_1) - D^2f(x_1) w) }$
\item{10.} \codent {\tt if} $\bigl(t + 3Df(x_0) + D^2f(x_0)w < 0 \bigr)$
  {\tt then return FALSE endif}
\item{11.} \codent {\tt if} $\bigl(60z - w\bigl(24Df(x_0) + 32Df(x_1) - 2t
  + w(3D^2f(x_0) - 5D^2f(x_1))\bigr) < 0\bigr)$
\item{}   \codent \codent {\tt then return FALSE endif}
\item{12.} \codent {\tt return TRUE}
\item{13.} {\tt endif}

\item{} {The following code considers the full quintic monotonicity
case studied by Ulrich and Watson [1994]. Given $\tau_1$, $\alpha$,
$\beta$, and $\gamma$ as defined by Ulrich and Watson, a quintic
piece is proven to be monotone if and only if
$\tau_1 > 0$, and $\alpha, \gamma > -(\beta+2)/2$ when $\beta \leq 6$,
and $\alpha, \gamma > -2 \sqrt{\beta-2}$ when $\beta > 6$.
Step 14 checks $\tau_1 \le 0$, Steps 19 and 20 determine monotonicity based
on $\alpha$, $\beta$, and $\gamma$.}

\item{14.} {\tt if} $\bigl( w\bigl(2\sqrt{Df(x_0)\,Df(x_1)} - 3(Df(x_0) +
  Df(x_1))\bigr) - 24z \leq 0 \bigr)$
\item{} \codent {\tt then return FALSE endif}
\item{15.} $t \: \bigl(Df(x_0)\, Df(x_1)\bigr)^{3/4}$
\item{16.} $\alpha \: (4 Df(x_1) - D^2f(x_1)w) \sqrt{Df(x_0)} / t$
\item{17.} $\gamma \: (4 Df(x_0) - D^2f(x_0)w) \sqrt{Df(x_1)} / t$
\item{18.} $\displaystyle \beta \: {60z/w + 3\bigl(w(D^2f(x_1) -
  D^2f(x_0)) - 8(Df(x_0) + Df(x_1))\bigr) \over 2 \sqrt{Df(x_0)\,Df(x_1)}}$
\item{19.} {\tt if} $(\beta \leq 6)$ {\tt then return}
$\bigl( \min\{\alpha,\gamma\} > - (\beta + 2) / 2 \bigr)$
\item{20.} {\tt else} {\tt return}
$\bigl( \min\{\alpha,\gamma\} > -2 \sqrt{\beta - 2}\,\bigr)$
\item{21.} {\tt endif}
}
\vskip 5mm
%% ----------------------------------------------------------------------

It is shown by Ulrich and Watson [1994] that when $0 = DQ(X_i) =
DQ(X_{i+1}) = D^2Q(X_i) = D^2Q(X_{i+1})$, the quintic polynomial
over $[X_i$, $X_{i+1}]$ is guaranteed to be monotone. Using this fact, the
following algorithm shrinks (in magnitude) initial derivative estimates
until a monotone spline is achieved and outlines the core routine in the
accompanying package.

%% ===================================================================
%%       Algorithm 3: MQSI - Monotone quintic spline interpolant
\vskip 3mm
{\parindent 0mm
{\bf Algorithm 3:} {\tt MQSI}$\bigl(X(1{:}n), Y(1{:}n) \bigr)$

where $(X_i,Y_i) \in \bbb{R}\times\bbb{R}$, $i = 1$, $\ldots$, $n$ are data
points. Returns monotone quintic spline interpolant $Q(x)$ such that
$Q(X_i) = Y_i$ and is monotone increasing (decreasing) on all
intervals that $Y_i$ is monotone increasing (decreasing).

}
{\parindent=6mm
\item{} Approximate first and second derivatives at $X_i$ with {\tt
  QUADRATIC\_FACET}.
\item{} {\tt for} $i\:1$ {\tt step} 1 {\tt until} $n$ {\tt do}
\item{} \codent $(u_i$, $v_i)\:$ {\tt QUADRATIC\_FACET}$(X$, $Y$, $i)$
\item{} {\tt enddo}
\item{} Identify and store all intervals where $Q$ is nonmonotone in a
queue $\cal Q$.
\item{} {\tt for} $i\:1$ {\tt step} 1 {\tt until} $n-1$ {\tt do}
\item{} \codent {\tt if not IS\_MONOTONE}$(X_i$, $X_{i+1}$, $Y_i$, $u_i$,
$v_i$, $Y_{i+1}$, $u_{i+1}$, $v_{i+1})$ {\tt then}
\item{} \codent\codent Add interval $\bigl(X_i$, $X_{i+1}\bigr)$ to
queue $\cal Q$.
\item{} \codent {\tt endif}
\item{} {\tt enddo}
\item{} {\tt do while} $\bigl($ queue $\cal Q$ of intervals is nonempty $\bigr)$
\itemitem{} Shrink (in magnitude) $DQ$ (in $u$) and $D^2Q$ (in $v$)
  that border intervals where $Q$ is nonmonotone.  
\itemitem{} Identify and store remaining intervals where $Q$ is
  nonmonotone in queue $\cal Q$.
\item{} {\tt enddo}
\item{} Construct and return a $B$-spline representation of $Q(x)$.
}
\vskip 3mm
%% ----------------------------------------------------------------------

Since {\tt IS\_MONOTONE} can handle both nondecreasing and
nonincreasing simultaneously by taking into account the sign of $z$,
Algorithm 3 produces $Q(x)$ that is monotone increasing (decreasing)
over exactly the same intervals that the data $(X_i$, $Y_i)$ is
monotone increasing (decreasing).

Given the minimum magnitude curvature nature of the initial derivative
estimates, it is desirable to make the smallest necessary changes to
the initial interpolating spline $Q$ while enforcing monotonicity. In
practice a binary search for the boundary of monotonicity is used in
place of solely shrinking $DQ$ and $D^2Q$ at breakpoints adjoining
{\it active\/} intervals, or intervals over which $Q$ is nonmonotone at
least once during the search. The binary search considers a Boolean
function $B_i(s)$, for $0 \le s \le 1$, that is true if the order six
polynomial $f(x)$ on $[X_i, X_{i+1}]$ matching derivatives
$f(X_i)=Y_i$, $Df(X_i)=s \,u_i$, $D^2f(X_i)=s \,v_i$ at $X_i$, and derivatives
$f(X_{i+1})=Y_{i+1}$, $Df(X_{i+1})=s \,u_{i+1}$, $D^2f(X_{i+1})=s \,v_{i+1}$
at $X_{i+1}$ is monotone, and false otherwise. As is outlined in Algorithm
3, the binary search is only applied at those breakpoints adjoining intervals
$[X_i, X_{i+1}]$ over which $Q$ is nonmonotone and hence $B_i(1)$ is false.
It is further assumed that there exists $0 \le s^* \le 1$ such that $B_i(s)$
is true for $0 \le s \le s^*$ and false for some $1 > s > s^*$. Since the
derivative conditions at interior breakpoints are shared by intervals left
and right of the breakpoint, the binary search is performed at all
breakpoints simultaneously.  Specifically, the monotonicity of $Q$ is
checked on all active intervals in each step of the binary search to
determine the next derivative modification at each breakpoint. The goal
of this search is to converge on the boundary of the monotone region in
the $(\tau_1$, $\alpha$, $\beta$, $\gamma)$ space (described in Ulrich and
Watson [1994]) for all intervals. This multiple-interval binary search
allows the value zero to be obtained for all (first and second) derivative
values in a fixed number of computations, hence has no effect on
computational complexity order. This binary search algorithm is outlined
below.

%% ===================================================================
%%              Algorithm 4: Global binary search
\vskip 5mm {\parindent 0mm {\bf Algorithm 4:} {\tt
    BINARY\_SEARCH}$\bigl(X(1{:}n), Y(1{:}n), u(1{:}n), v(1{:}n) \bigr)$

where $(X_i,Y_i) \in \bbb{R}\times\bbb{R}$, $i = 1$, $\ldots$, $n$ are
data points, and $Q(x)$ is a quintic spline interpolant such that
$Q(X_i) = Y_i$, $DQ(X_i) = u_i$, $D^2Q(X_i) = v_i$. Modifies
derivative values ($u$ and $v$) of $Q$ at data points to ensure
{\tt IS\_MONOTONE} is true for all intervals defined by adjacent data
points, given a desired precision $\mu \in {\bbb R}$. proj$\bigl(w$,
int$(a,b)\bigr)$ denotes the projection of $w$ onto the closed
interval with endpoints $a$ and $b$.

}
{\parindent=6mm
\item{} Initialize the step size $s$, make a copy of data defining $Q$,
and construct three queues necessary for the multiple-interval binary search.
\item{} $s \: 1$
\item{} $(\hat u,\hat v) \: (u,v)$
\item{} {\tt searching} $\:$ {\tt TRUE}
\item{} {\tt checking} $\:$ empty queue for holding left indices of {\it intervals}
\item{} {\tt growing} $\:$ empty queue for holding indices of {\it data points}
\item{} {\tt shrinking} $\:$ empty queue for holding indices of {\it data points}
\item{} {\tt for} $i\:1$ {\tt step} 1 {\tt until} $n-1$ {\tt do}
\itemitem{} {\tt if not IS\_MONOTONE}$\bigl( X_i$, $X_{i+1}$, $Y_i$, $u_i$,
$v_i$, $Y_{i+1}$, $u_{i+1}$, $v_{i+1} \bigr)$ {\tt then}
\itemitem{} \codent Add data indices $i$ and $i+1$ to queue {\tt shrinking}.
\itemitem{} {\tt endif}
\item{} {\tt enddo}
\item{} {\tt do while} $\bigl(${\tt searching or} ({\tt shrinking} is
nonempty)$\bigr)$
\itemitem{} Compute the {\it step size} $s$ for this iteration of the search.
\itemitem{} {\tt if  searching  then} $s \: \max\{\mu, s/2\}$
{\tt else} $s \: 3s/2$ {\tt endif}
\itemitem{} {\tt if} $(s=\mu)$ {\tt then searching} $\:$ {\tt FALSE};
  clear queue {\tt growing endif}
\itemitem{} Increase in magnitude $u_i$ and $v_i$ for all data indices
  $i$ in {\tt growing} such that the points $X_i$ are strictly adjoining
  intervals over which $Q$ is monotone.
\itemitem{} {\tt for} ($i \in$ {\tt growing}) {\tt and} ($i \not\in$
{\tt shrinking}) {\tt do}
\itemitem{} \codent $u_i \: \hbox{proj}\bigl(u_i + s\,\hat u_i, \hbox{int}
  (0,\hat u_i)\bigr)$
\itemitem{} \codent $v_i \: \hbox{proj}\bigl(v_i + s\,\hat v_i, \hbox{int}
  (0,\hat v_i)\bigr)$
\itemitem{} \codent Add data indices $i-1$ (if not 0) and $i$ (if not $n$) to queue
  {\tt checking}.
\itemitem{} {\tt enddo}
\itemitem{} Decrease in magnitude $u_i$ and $v_i$ for all data indices
  $i$ in {\tt shrinking} and ensure those data point indices are placed
  into {\tt growing} when {\tt searching}.
\itemitem{} {\tt for} $i \in$ {\tt shrinking do}
\itemitem{} \codent If {\tt searching}, then add index $i$ to queue
  {\tt growing} if not already present.
\itemitem{} \codent $u_i \: \hbox{proj}\bigl(u_i - s\,\hat u_i, \hbox{int}
  (0,\hat u_i)\bigr)$
\itemitem{} \codent $v_i \: \hbox{proj}\bigl(v_i - s\,\hat v_i, \hbox{int}
  (0,\hat v_i)\bigr)$
\itemitem{} \codent Add data indices $i-1$ (if not 0) and $i$ (if not $n$) to queue
  {\tt checking}.
\itemitem{} {\tt enddo}
\itemitem{} Empty queue {\tt shrinking}, then check all intervals left-indexed
  in queue {\tt checking} for monotonicity with {\tt IS\_MONOTONE},
  placing data endpoint indices of intervals over which $Q$ is nonmonotone
  into queue {\tt shrinking}.
\itemitem{} Clear queue {\tt shrinking}.
\itemitem{} {\tt for} $i \in$ {\tt checking do}
\itemitem{} \codent
  {\tt if not IS\_MONOTONE}$\bigl( X_i$, $X_{i+1}$, $Y_i$, $u_i$,
  $v_i$, $Y_{i+1}$, $u_{i+1}$, $v_{i+1} \bigr)$ {\tt then}
\itemitem{} \codent \codent Add data indices $i$ and $i+1$ to {\tt shrinking}.
\itemitem{} \codent {\tt endif}
\itemitem{} {\tt enddo}
\itemitem{} Clear queue {\tt checking}.
\item{} {\tt enddo }
}
\vskip 5mm
%% ----------------------------------------------------------------------

%% \itemitem{} Take search steps either modifying derivative values
%%   towards monotonicity (zero) or towards the original values over all
%%   intervals that are either nonmonotone or previously modified.

In the subroutine {\tt MQSI}, $\mu = 2^{-26}$, which results in 26
guaranteed search steps for all intervals that are initially
nonmonotone. An additional 43 steps could be required to reduce a
derivative magnitude to zero with step size growth rate of $3/2$. This
can only happen when $Q$ becomes nonmonotone on an interval for the
first time while the step size equals $\mu$, but for which the only
viable solution is a derivative value of zero. The maximum number of
steps is due to the fact that $\sum_{i=0}^{42} \mu (3/2)^i > 1$. In
total {\tt BINARY\_SEARCH} search could require 69 steps.