### 2020-07-28 3-Disseration/error.tex

This chapter presents theoretical results bounding the error of
(piecewise) linear interpolation. The error analysis relies on linear
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
paper compute predictions as convex combinations of observed function
values, which may allow for straightforward extensions of this error
bound.

\begin{plainlemma}
  \label{lemma:1}
  Let $S \subset \mathbb{R}^d$ be open and convex, $f: S \rightarrow
  \mathbb{R}$, and let $\nabla f(x)$ be $\gamma$-Lipschitz continuous
  in the $2$-norm. Then for all $x,y \in S$
  $$\big|f(y) - f(x) - \langle \nabla f(x), y - x \rangle \big| \leq \frac{\gamma \|y - x\|_2^2}{2}.$$
\end{plainlemma}

\begin{proofdot}
  Consider the function $g(t) = f \big((1-t) x + t y \big)$, $0 \leq t
  \leq 1$, whose derivative $g'(t) = \big\langle \nabla f \big((1-t) x
  + t y \big), y - x \big\rangle$ is the directional derivative of $f$
  in the direction $(y - x).$
  \begin{align*}
    \big|f(y) - f(x) - \langle &\nabla f(x), y - x \rangle \big|
        = \big|g(1) - g(0) - g'(0) \big| \\
       &= \bigg| \int_0^1 g'(t) - g'(0)\ dt \bigg| \leq \int_0^1 \big|g'(t) - g'(0)\big|\ dt \\
       &= \int_0^1 \bigg| \big \langle \nabla f\big((1-t)x + ty\big) - \nabla f(x), y - x \big \rangle \bigg|\ dt \\
       &\leq \int_0^1 \big \| \nabla f\big((1-t)x + ty\big) - \nabla f(x) \big \|_2\ \| y - x \|_2\ dt \\
       &\leq \int_0^1 \big ( \gamma\ \|y-x\|_2 \big) \ \big( \|y-x\|_2 \big) t\ dt = \frac{\gamma \|y - x\|_2^2}{2}.
  \end{align*}
  \qed
\end{proofdot}

\hfill

\begin{plainlemma}
  \label{lemma:2}
  Let $x, y, v_i \in \mathbb{R}^d$, $c_i \in \mathbb{R}$, and
  $|\langle y - x, v_i \rangle| \leq c_i$ for $i = 1$, $\ldots$, $d.$
  If $M = (v_1$, $\ldots$, $v_d)$ is nonsingular, then
  $$\|y - x\|_2^2 \leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2,$$
  where $\sigma_d$ is the smallest singular value of $M.$
\end{plainlemma}

\begin{proofdot}
  Using the facts that $M$ and $M^t$ have the same singular values,
  and $\|M^tw\|_2 \geq \sigma_d \|w\|_2$, gives
  \begin{align*}
    \|y - x\|_2^2 &\leq \frac{\|M^t (y - x)\|_2^2}{\sigma_d^2} \\
                  &=    \frac{1}{\sigma_d^2} \sum_{i=1}^d \langle y - x, v_i \rangle^2 \\
                  &\leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2.
  \end{align*}
  \qed
\end{proofdot}

%% \newpage

\begin{plainlemma}
  \label{lemma:3}
  Given $f$, $\gamma$, $S$ as in {\it Lemma \ref{lemma:1}}, let $X =
  \{x_0$, $x_1$, $\ldots$, $x_d\}$ $\subset S$ be the vertices of a
  $d$-simplex, and let $\hat f(x) = \langle c, x - x_0 \rangle +
  f(x_0)$, $c \in \mathbb{R}^d$ be the linear function interpolating
  $f$ on $X.$ Let $\sigma_d$ be the smallest singular value of the
  matrix $M = (x_1 - x_0$, $\ldots$, $x_d - x_0)$, and $k =
  \max\limits_{1\ \leq\ j\ \leq\ d} \|x_j - x_0\|_2.$ Then
  $$\big\|\nabla f(x_0) - c\big\|_2 \leq \frac{\sqrt{d} \, \gamma k^2}{2 \sigma_d}.$$
\end{plainlemma}

\begin{proofdot}
  \def\g{\gamma_g}
  Consider $g(t)=f(x(t)) - \hat f(x(t))$ along
  the line segment $x(t)=(1-t)x_0 + tx_j$, $0\le t\le1$, from $x_0$ to
  $x_j$.  Observe that $g(0)=f(x_0) - \hat f(x_0) = 0$, $g(1)=f(x_j) -
  \hat f(x_j) = 0$, and $g'(t)$ is $\g$-Lipschitz continuous with $\g
  = \gamma \|x_j-x_0\|_2^2$. The following is visualized in 
  Figure \ref{fig:lemma3}.

  Suppose $g'(0) > \g/2$. Then $|g'(0)-g'(t)| \le \g t \implies g'(0)
  - \g t \le g'(t)$, and the line $w=g'(0)-\g t$ intersects the
  $t$-axis at $\tilde t=g'(0)/\g > 1/2$. $\tilde t<1$ necessarily,
  since by Rolle's Theorem there exists $0<z<1$ such that $g'(z)=0$
  and so $g'(0) - \g z \le g'(z) = 0$. Now by integrating
  $$\frac{g'(0)^2}{2\g} = \frac{g'(0)\tilde t}{2} = \int_0^{\tilde t}
  (g'(0)-\g t)dt \le \int_0^{\tilde t} g'(t)dt = g(\tilde t),$$ and
  using $g'(0) > \g / 2$
  \begin{align*}
    \frac{-g'(0)^2}{2\g} < g'(0) - \frac{g'(0)^2}{2\g} - \g/2 =
    \frac{(1-\tilde t)(g'(0)-\g)}{2} &= \int_{\tilde t}^1 (g'(0)-\g
    t)dt \\ &\le \int_{\tilde t}^1 g'(t)dt = -g(\tilde t),
  \end{align*}
  \noindent a contradiction for the value of $g(\tilde t)$. A similar
  contradiction arises for $g'(0)<-\g/2$. Therefore $|g'(0)| \le
  \g/2$. In terms of $f$,
  $$\bigl|\langle \nabla f(x_0)-c,x_j-x_0\rangle\bigr| = |g'(0)| \le
  \gamma \|x_j-x_0\|_2^2/2 \le \gamma k^2 / 2,$$ which holds for all
  $1\le j\le d$.  Finally, using {\it Lemma \ref{lemma:2}},
  $$\|\nabla f(x_0)-c\|_2^2 \le \frac{d}{\sigma_d^2}(\gamma k^2/2)^2
  \implies \|\nabla f(x_0)-c\|_2 \le \frac{\sqrt{d} \, \gamma
    k^2}{2\sigma_d}.$$ \qed
\end{proofdot}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth,]{Figures/NA/example_lemma_3.pdf}
  \def\g{\gamma_g}
  \caption{Three different scenarios visualizing {\it
      Lemma \ref{lemma:3}}, where $g(t)$ is the difference between a
    piecewise linear interpolant and the approximated function along a
    normalized line segment between interpolation points, $g'(t)$ is
    $\g$-Lipschitz continuous, and $w$ and $\tilde t$ are defined in
    the proof. Leftmost is a randomly chosen permissible shape of $g$
    and $g'$. The middle is the only possible shape of $g$ and $g'$
    given $g'(0) = \g / 2$, establishing the case of equality in the
    lemma. Rightmost is the resulting contradiction when $g'(0) > \g /
    2$, notice it is impossible to ensure $g'(t)$ is $\g$-Lipschitz
    continuous and satisfy $g(1) = 0$ (highlighted with red circle on
    the right).
  \vspace{-.1cm}}
  \label{fig:lemma3}
\end{figure}

\begin{onetheorem}
  Under the assumptions of {\it Lemma \ref{lemma:1}} and {\it Lemma
    \ref{lemma:3}}, for $z \in S$,
  $$ \big|f(z) - \hat f(z)\big| \leq \frac{\gamma \|z - x_0\|_2^2}{2} + \frac{\sqrt{d} \, \gamma k^2}{2 \sigma_d} \|z - x_0\|_2.$$
\end{onetheorem}

\begin{proofdot}
  Let $v = \nabla f(x_0) - c$, where $\|v\|_2 \leq \sqrt{d}\, \gamma k^2 / (2 \sigma_d)$ by {\it Lemma \ref{lemma:3}}. Now
  \begin{align*}
    \big|f(z) - \hat f(z)\big|
       &= \big|f(z) - f(x_0) - \langle c, z - x_0 \rangle \big| \\ 
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) - v, z - x_0 \rangle \big| \\
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle + \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \big| \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \|v\|_2 \|z - x_0\|_2 \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0), z - x_0 \rangle \big| + \big(\sqrt{d}\,\gamma k^2 / (2 \sigma_d)\big) \|z - x_0\|_2 \\
       &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \frac{\sqrt{d}\,\gamma k^2}{2 \sigma_d} \|z - x_0\|_2,
  \end{align*}
  where the last inequality follows from {\it Lemma \ref{lemma:1}}.
  %
  \qed
\end{proofdot}

In summary, the approximation error of a linear (simplicial)
interpolant tends quadratically towards zero when approaching observed
data only when the diameter of the simplex is also reduced at a
proportional rate. Only linear convergence to the true function
is achieved in practice without the
incorporation of additional observations, by moving closer to
  interpolation points. Notice that the approximation error is
largely determined by the spacing of observed data. Predictions made
by simplices whose vertices are not well-spaced (i.e., have large
diameter, or are nearly contained in a hyperplane) have higher error.

That the theoretical error bound is sharp can be observed with the
test function $q(x) = \|x\|_2^2,$ $x \in \mathbb{R}^d,$ and the
simplex defined by vertices $X = \{0, e_1, \ldots, e_d\},$ where $e_i$
is the $i$-th standard basis vector in $\mathbb{R}^d,$ $e = (1$,
$\ldots$, $1) \in \mathbb{R}^d,$ and $0$ denotes the zero vector in
any dimension.  The constants relevant to the error bound are
$$ \gamma = 2, \qquad \sigma_d = 1, \qquad k = 1, \qquad x_0 = 0,
\qquad \hat q(x) = \langle e, x - 0\rangle + q(0). $$
\noindent Noting that $q(0) = 0,$ the approximation error at $z =
-(1/2)e$ is
$$ \big|q(z) - \hat q(z)\big| = \big|\|z\|_2^2 - \langle e,z
\rangle\big| = \big| d/4 + d/2 \big| = 3d/4, $$
\noindent while the error bound from the theorem gives

\vspace{-2mm}
\begin{align*}
  \big|q(z) - \hat q(z)\big|
  &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \frac{\sqrt{d}\,\gamma k}
      {2 \sigma_d} \|z - x_0\|_2 \\
  &{}=  \|z\|_2^2 + \sqrt{d} \|z\|_2 \\
  &{}=  d/4 + d/2 = 3d/4.
\end{align*}
\vspace{-2mm}

\noindent Acknowledging that the error bound is sharp, it may be of
interest to observe the error of piecewise linear approximation
techniques on an analytic test function.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.9\textwidth,]{Figures/NA/Fig2.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Number of Data Points};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Absolute Error};
  \end{tikzpicture}
  \caption{Delaunay and MLP approximations are constructed from Fekete
    points over the unit cube evaluating the test function $f(x) =
    \cos(\|x\|_2)$ for $x \in \mathbb{R}^2$. The figure shows the
    first/third quartiles at the box bottom/top, the second quartile
    (median) at the white bar, median 95\% confidence interval (cones,
    barely visible in figure), and whiskers at 3/2 of the adjacent
    interquartile ranges, for the absolute prediction error for each
    model at $1000$ random evaluation points. The left plot observes a
    perfect interpolation problem with exact evaluations of $f.$ The
    right plot observes a regression problem with uniform random noise
    giving values in $[.9 f(x),\ 1.1f(x)]$ for each $x.$ Both axes are
    log scaled.}
  \label{fig:convergence-2d}
\end{figure}

%% ===================================================================
\vspace{-2mm}
\section{Demonstration on an Analytic Test Function}
\label{sec:analytic}

The theoretical results constructed in Section \ref{sec:theory} for
(piecewise) linear interpolation are promising and apply directly to
Delaunay interpolation, however they are difficult to interpret in
context with approximation algorithms that do not have similar known
uniform error bounds. For that reason, an analytic function is used to
measure the error of a piecewise linear interpolant (Delaunay) and a
piecewise linear regressor (the MLP) when provided an increasing
amount of data with varying levels of random noise. Only
Delaunay and the MLP are considered in this demonstration because
all other mentioned techniques are not strictly piecewise linear
approximations.

The test function chosen here for analysis resembles the
\textit{oscillatory} function used by \cite{barthelmann2000high}.
However, a slight modification is made to remove the simple dot
product structure (which is favorable for the MLP). Let $f(x)
=\cos(\|x\|_2)$ for $x \in \mathbb{R}^d$ where $d$ is either $2$
(Fig. \ref{fig:convergence-2d}) or $20$
(Fig. \ref{fig:convergence-20d}). This function has a bounded change
in gradient $2$-norm and hence meets the necessary Lipschitz condition
for the error bound. Data points and approximation points for this
test will be within the unit cube $[0,1]^d$.

The error of a linear interpolant constructed from well spaced points
is dominated by the distance to the nearest data point. In order to
uniformly decrease the distance to the nearest data point across the
unit cube, exponentially more data points must be available for
approximation. For this experiment $N = 2(4^i),$ $i = 1,$ $\ldots,$
$6,$ chosen points are kept well spaced by computing approximate
Fekete points. The following experiment was also run with a
  Latin hypercube design \cite{park1994optimal}, which produced almost
  identical results that are not reported here. Fekete points have
a history in potential theory \cite{kovari_pommerenke_1968} and are
most generally defined as those points that maximize the absolute
value of a Vandermonde determinant. Here, the QR method outlined in
\cite{bos2010computing} is used to identify approximate Fekete points
from a Vandermonde matrix. A multivariate polynomial basis of size
$2N$ is constructed containing all polynomials of degree $\leq n$ for
the largest $n$ such that ${d+n \choose n} \leq 2N$ and $2N - {d+n
  \choose n}$ arbitrarily selected polynomials of degree $n + 1.$ The
Vandermonde matrix for these $2N$ polynomials is used to select $N$
approximate Fekete points.

Finally an additional aspect is added to this test problem by
incorporating random uniform noise into the evaluations of $f.$ For
each test two experiments are executed, one with exact function
evaluations (an interpolation problem) and one with a
constant signal-to-noise ratio (SNR) of 10:1 (a regression
  problem).

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.9\textwidth,]{Figures/NA/Fig3.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Number of Data Points};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Absolute Error};
  \end{tikzpicture}
  \caption{Delaunay and MLP approximations are constructed from Fekete
    points over the unit cube evaluating the test function $f(x) =
    \cos(\|x\|_2)$ for $x \in \mathbb{R}^{20}$. The details are the
    same as for Fig. \ref{fig:convergence-2d}.}
  \label{fig:convergence-20d}
\end{figure}

The bound from the theorem suggests that by increasing the number of
well-spaced data points, approximation error can be reliably
decreased. The $d = 2$ test seen on the left half of Figure
\ref{fig:convergence-2d} shows a consistent decrease in error for
Delaunay and also shows the eventual accuracy plateau obtained by a
parametric regression form (the MLP, at roughly 500 points). On the
right hand side of Figure \ref{fig:convergence-2d} the random noise
clearly prohibits Delaunay from converging to $f$, while the MLP is
able to improve its approximation with more data points on average.
The convergence result for a very low-dimensional problem like this is
expected.  However, intuition fails for higher dimensional problems.

Figure \ref{fig:convergence-20d} shows the test function with $d = 20$
presents a significantly more challenging approximation problem than
its counterpart in low dimension. The same increase in number of data
points from 32 to 8192 causes no apparent improvement in approximation
for either the noise-free or the noisy problems. Perhaps unexpectedly,
the interpolation technique performs better than the regression
technique on the noisy data (right) in Figure
\ref{fig:convergence-20d}, and worse than the regression technique on
the noise-free data (left). This result emphasizes the relevance of
interpolation for problems in high dimension. It also reveals that the
outcome of MLP regressions can get worse when adding more data points.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.9\textwidth,]{Figures/NA/theorem-convergence.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Number of Data Points};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Error, Distance, and Smallest SV };
  \end{tikzpicture}
  \caption{The distribution of absolute error, distance to
    the nearest data point, smallest singular value (SV) and the
    longest edge of the simplex containing each approximation point in
    the tests from Fig. \ref{fig:convergence-2d} and
    Fig. \ref{fig:convergence-20d} for Delaunay. In two dimensions it
    can be seen that $\|z - x_0\|_2$, $\sigma_d$, and $k$ all shrink
    at the same rate for well-spaced approximation points, resulting
    in a faster rate of decrease for approximation error. Notice that
    in higher dimension the data remains sparse even with thousands of
    data points, and the decay in data spacing is more prominant. The
    relatively small reduction in $k$ along with the decrease in
    $\sigma_d$ explain the minimal reduction in error seen by Delaunay
    in Fig. \ref{fig:convergence-20d}.}
  \label{fig:data-spacing}
\end{figure}

The results achieved for both $d = 2$ and $d = 20$ align with the
theoretical error bound as can be seen in Figure
\ref{fig:data-spacing}. In low dimension, thousands of data points
meaningfully reduce the sparsity of the approximation problem.
However, in higher dimension it takes many more points to achieve a
reduction in data sparsity while simultaneously being more difficult
to produce well-conditioned simplices. This evidences the
inherent challenge of data sparsity in high dimension approximation
problems. The analytic results presented are not caused by the chosen
test function, but rather the exponential increase in complexity that
accompanies increased dimension.

In summary, the regime of signal-to-noise ratios that result in
competitively accurate interpolants is greater for moderate to high
dimensional problems due to increased sparsity. Acknowledging the
viability of interpolation for problems of moderate dimension, the
next section will consider real-world problems of similar proportion
(thousands of examples in tens of dimensions).

%% ===================================================================
\vspace{-2mm}
\section{Data and Empirical Analysis}
\label{sec:error_data}

This section extends the comparison of interpolation and regression
algorithms to a sample of real-world problems. Five different data
sets of varying dimension and application are utilized to construct
approximations and compare the accuracy of different techniques.

In the following five subsections the sources and targets of each test
data set are described, as well as challenges and limitations related
to approximating these data. The distribution of response values being
modeled is presented followed by the distribution of approximation
errors for each algorithm. The plots for all five data sets have the
same format.

All five data sets are rescaled such that the domain of approximation
is the unit hypercube. The range of the first four data sets is the
real numbers, while the range of the fifth data set is the space of
cumulative distribution functions. All approximation techniques are
applied to the first four data sets, while only those interpolants
whose approximations are convex combinations of observed data are
applied to the final data set.

All approximations are constructed using $k$-fold cross validation as
described in \cite{kohavi1995study} with $k=10$. This approach
randomly partitions data into $k$ (nearly) equal sized sets. Each
algorithm is then evaluated by constructing an approximation over each
unique union of $k-1$ elements of the partition, making predictions
for points in the remaining element. As a result, each observed data
point is used in the construction of $k-1$ different approximations
and is approximated exactly once. The $k$-fold cross validation method
is data-efficient and provides an unbiased estimate of the expected
prediction error \cite{kohavi1995study}, however it should be noted
that neither this method nor others can provide a universally unbiased
estimator for the variance of prediction error \cite{bengio2004no}.

In addition to the figures displaying approximation results for each
data set, scatter plots of predicted versus actual values
  and tables of accompanying numerical results including
  timings and quartiles of prediction error are located in the
Appendix (Section \ref{sec:appendix}). All of the test data sets
capture underlying functions that are almost certainly stochastic. As
described in Section \ref{sec:introduction}, regression techniques
appear most appropriate for these problems. However, typically data
grows exponentially more sparse with increasing dimension. Given that
sparse data regressors tend towards interpolation and, as demonstrated
in Section \ref{sec:analytic}, interpolants produce similar (if not
identical) results, it is presumed that interpolants are equally
viable approximation techniques on these problems.

%% -------------------------------------------------------------------
\subsection{Forest Fire ($n = 504$, $d = 12$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/NA/Fig4.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Forest Fire Area Burned};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of forest fire area burned under recorded weather
    conditions. The data is presented on a $\ln$ scale because
    most values are small with exponentially fewer fires on record
    that burn large areas.}
  \label{fig:hist-forest-fire}

  \vspace{.3cm}

  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/NA/Fig5.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Area Burned Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of area
    that would be burned given environment conditions. $10$-fold cross
    validation as described in the beginning of Section \ref{sec:error_data}
    is used to evaluated each algorithm. This results in exactly one
    prediction from each algorithm for each data point. These boxes
    depict the median (middle bar), median $95\%$ confidence interval
    (cones), quartiles (box edges), fences at $3/2$ interquartile
    range (whiskers), and outliers (dots) of absolute prediction error
    for each model. Similar to Figure \ref{fig:hist-forest-fire}, the
    errors are presented on a $\ln$ scale. The numerical data
    corresponding to this figure is provided in Table
    \ref{table:error-forest-fire} in the Appendix.}
  \label{fig:error-forest-fire}
\end{figure}

The forest fire data set \cite{cortez2007data} describes the area of
Montesinho park burned over months of the year along with
environmental conditions. The twelve dimensions being used to model
burn area are the $x$ and $y$ spatial coordinates of burns in the
park, month of year (mapped to $x$, $y$ coordinates on a unit circle),
the FFMC, DMC, DC, and ISI indices (see source for details), the
temperature, relative humidity, wind speed, and outdoor rain. The
original analysis of this data set demonstrated it to be difficult to
model, likely due to the skew in response values.

As suggested by Figure \ref{fig:error-forest-fire}, the SVR has the
lowest absolute prediction errors for $80\%$ of the data, with MLP and
Delaunay being the nearest overall competitors. The effectiveness of
SVR on this data suggests the underlying function can be defined by
relatively few parameters, as well as the importance of capturing the
low-burn-area data points.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Parkinson's Telemonitoring ($n = 5875$, $d = 19$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/NA/Fig6.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Parkinson's Total UPDRS Score};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of the Parkinson's patient total UPDRS clinical
    scores that will be approximated by each algorithm.}
  \label{fig:hist-parkinsons}

  \vspace{.3cm}

  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/NA/Fig7.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Total UPDRS Score Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the total UPDRS score
    given audio features from patients' home life, using $10$-fold
    cross validation. These boxes depict the median (middle bar),
    median $95\%$ confidence interval (cones), quartiles (box edges),
    fences at $3/2$ interquartile range (whiskers), and outliers
    (dots) of absolute prediction error for each model. The numerical
    data corresponding to this figure is provided in Table
    \ref{table:error-parkinsons} in the Appendix.}
  \label{fig:error-parkinsons}
\end{figure}

The second data set for evaluation \cite{tsanas2010accurate} is
derived from a speech monitoring study with the intent to
automatically estimate Parkinson's disease symptom development in
Parkinson's patients. The function to be approximated is a
time-consuming clinical evaluation measure referred to as the UPDRS
score. The total UPDRS score given by a clinical evaluation is
estimated through 19 real numbers generated from biomedical voice
measures of in-home sound recordings.

Figure \ref{fig:error-parkinsons} shows the ShepMod algorithm has the
lowest minimum, first quartile, and median of absolute errors for this
problem, while providing the best prediction $66\%$ of the time. The
MLP has the lowest third quartile and provides the best prediction for
$32\%$ of approximations. The dominance of ShepMod may be due in part
to regular-interval total UPDRS scores provided by clinicians,
favoring a nearest-neighbor strategy of prediction.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Australian Daily Rainfall Volume ($n = 2609$, $d = 23$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/NA/Fig8.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Daily Rainfall in Sydney};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of daily rainfall in Sydney, Australia, presented
    on a $\ln$ scale because the frequency of larger amounts of
    rainfall is significantly less. There is a peak in occurrence of
    the value $0$, which has a notable effect on the resulting model
    performance.}
  \label{fig:hist-weather}

  \vspace{.3cm}

  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/NA/Fig9.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Sydney Rainfall Tomorrow Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of
    rainfall expected on the next calendar day given various sources
    of local meteorological data, using $10$-fold cross validation.
    These boxes depict the median (middle bar), median $95\%$
    confidence interval (cones), quartiles (box edges), fences at
    $3/2$ interquartile range (whiskers), and outliers (dots) of
    absolute prediction error for each model. The errors are presented
    on a $\ln$ scale, mimicking the presentation in Figure
    \ref{fig:hist-weather}. The numerical data corresponding to this
    figure is provided in Table \ref{table:error-weather} in the
    Appendix.}
  \label{fig:error-weather}
\end{figure}

The third data set for the total daily rainfall in Sydney, Australia
\cite{williams2009rattle} provides a slightly higher dimensional
challenge for the interpolants and regressors. This data is composed
of many meteorological readings from the region in a day including:
min and max temperatures, sunshine, wind speed directions (converted
to coordinates on a circle), wind speeds, and humidities throughout
the day, day of the year (converted to coordinates on a circle), and
the model must predict the amount of rainfall tomorrow.

While Figure \ref{fig:hist-weather} makes MARS look far better than
other techniques, it only provides the best prediction for $11\%$ of
points. The MLP has the lowest absolute error for $56\%$ of points and
LSHEP is best for $28\%$. MARS likely achieves such a low first
quartile due to the high occurrence of the value zero in the data.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Credit Card Transaction Amount ($n = 5562$, $d = 28$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/NA/Fig10.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Credit Card Transaction Amount};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of credit card transaction amounts, presented on
    a $\ln$ scale. The data contains a notable frequency peak around
    $\$1$ transactions. Fewer large purchases are made, but some large
    purchases are in excess of five orders of magnitude greater than
    the smallest purchases.}
  \label{fig:hist-credit-card}

  \vspace{.3cm}

  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/NA/Fig11.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Transaction Amount Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the expected
    transaction amount given transformed (and obfuscated) vendor and
    customer-descriptive features, using $10$-fold cross validation.
    These boxes depict the median (middle bar), median $95\%$
    confidence interval (cones), quartiles (box edges), fences at
    $3/2$ interquartile range (whiskers), and outliers (dots) of
    absolute prediction error for each model. The absolute errors in
    transaction amount predictions are presented on a $\ln$ scale,
    just as in Figure \ref{fig:hist-credit-card}. The numerical data
    corresponding to this figure is provided in Table
    \ref{table:error-credit-card} in the Appendix.}
  \label{fig:error-credit-card}
\end{figure}

The fourth test data set, and the final with a real-valued range, is a
collection of credit card transactions
\cite{pozzolo2015calibrating}. The provided data carries no direct
real-world meaning, being the output of a principle component analysis
on the original hidden source data. This obfuscation is done to
protect the information of the credit card users. This data has the
largest dimension of all considered, at 28. A model for this data
predicts the transaction amount given the vector of principle
component information.

As can be seen in Figure \ref{fig:error-credit-card}, the MLP
outperforms all other algorithms at the first, second, third, and
fourth quartiles. The MLP produces the lowest absolute error
prediction for $80\%$ of transactions, Delaunay bests the remaining
$20\%$. It is likely that with less data, Delaunay would be the best
performer.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{High Performance Computing I/O ($n = 3016$, $d = 4$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/NA/Fig12.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Read Throughput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of the raw throughput values recorded during all
    IOzone tests across all system configurations. The distribution is
    skewed right, with few tests having significantly higher
    throughput than most others. The data is presented on a $\ln$
    scale.}
  \label{fig:hist-throughput}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/NA/Fig13.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Read Throughput KS Statistic};
  \end{tikzpicture}
  \caption{The models directly capable of predicting distributions are
    applied to predicting the expected CDF of I/O throughput at a
    previously unseen system configuration, using $10$-fold cross
    validation. The KS statistic (max norm) between the observed
    distribution at each system configuration and the predicted
    distribution is recorded and presented above. Note that the above
    figure is \textit{not} log-scaled like Figure
    \ref{fig:hist-throughput}. The numerical data corresponding to
    this figure is provided in Table \ref{table:error-throughput} in
    the Appendix.}
  \label{fig:error-throughput}
\end{figure}

The final of five data sets is derived from \cite{cameron2019moana},
which provides four-dimensional distribution data by executing the
IOzone benchmark \cite{iozone} on a computer system and varying the
system's file size, record size, thread count, and CPU frequency. At
each configuration, IOzone samples the I/O file-read throughput (in
bytes per second) 150 times. Empirical distribution function points
are computed from each set of 150 executions, which are interpolated
with a piecewise cubic Hermite interpolating polynomial
\cite{fritsch1980monotone} to approximate the CDF. All interpolation
algorithms with the exception of LSHEP are used to approximate these
CDFs from system configurations.

Delaunay achieves the lowest KS statistic (max norm difference) for
$62\%$ of approximations, while Voronoi is best for the remaining
$38\%$. Figure \ref{fig:error-throughput} shows that while Delaunay
may have more best predictions, the behavior of Voronoi may be
preferable.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth]{Figures/NA/Fig14.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {KS Statistic for Predicted vs. Actual};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count of KS Statistic};
  \end{tikzpicture}
  \caption{Histograms of the prediction error for each interpolant
    that produces predictions as convex combinations of observed data,
    using $10$-fold cross validation. The histograms show the KS
    statistics for the predicted throughput distribution versus the
    actual throughput distribution. The four vertical lines represent
    cutoff KS statistics given $150$ samples for commonly used
    $p$-values 0.05, 0.01, 0.001, $10^{-6}$, respectively. All
    predictions to the right of a vertical line represent CDF
    predictions that are significantly different (by respective
    $p$-value) from the actual distribution according to the KS
    test. The numerical counterpart to this figure is presented in
    Table \ref{table:null-hypothesis-results}.}
  \label{fig:throughput-prediction}
\end{figure}

\begin{table}
  \renewcommand{\arraystretch}{1.3}
  \centering
  \begin{tabular}{c|c|c|c|c}
               & $p = .05$         & $p = .01$         & $p = .001$        & $p = 10^{-6}$\\
    \hline
    Delaunay   & $\mathbf{50.3}\%$ & $\mathit{43.5}\%$ & $\mathit{36.2}\%$ & $\mathit{24.7}\%$\\
    ShepMod    & $\mathit{51.4}\%$ & $44.8\%$          & $38.1\%$          & $27.7\%$\\
    Voronoi    & $52.6\%$          & $\mathbf{43.4}\%$ & $\mathbf{34.4}\%$ & $\mathbf{19.1}\%$\\
    BoxSplines & $99.4\%$          & $98.5\%$          & $96.6\%$          & $89.3\%$\\
  \end{tabular}
  \caption{Numerical counterpart of the histogram data presented in
    Figure \ref{fig:throughput-prediction}. The columns display the
    percent of null hypothesis rejections by the KS-test when provided
    different selections of $p$-values for each algorithm. The
    algorithm with the lowest rejection rate at each $p$ is boldface,
    while the second lowest is italicized.}
  \label{table:null-hypothesis-results}
\end{table}

%% ===================================================================

Figure \ref{fig:throughput-prediction} expands on the KS statistic
results presented in Figure \ref{fig:error-throughput}. Agglomerate
errors for each algorithm resemble a Gamma distribution. The
percentages of significant prediction errors with varying $p$-values
are on display in Table \ref{table:null-hypothesis-results}. When
considering the $p=0.001$ results for each technique, a little over
one third of the predicted CDFs are significantly different from the
measured (and presumed) correct CDFs. However, it should be noted that
with 150 samples, the error of an empirical distribution function
(EDF) can reasonably be upwards of $.1$, which serves as a rough
estimate for the lower limit of achievable error. Globally, only a
third of Voronoi predictions fail to capture \textit{all} of the
characteristics of the CDFs at new system configurations.

\section{Discussion of Empirical Results}
\label{sec:discussion}

\begin{table}
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l}
    Algorithm & \multicolumn{2}{c|}{Avg. \% Best} & \multicolumn{2}{c|}{\multilinecell{Avg. Fit or \\ Prep. Time (s)}} & \multicolumn{2}{c}{Avg. App. Time (s)}\\
    \hline
    MARS        & \quad\quad$4$&$5$          & \quad\quad$20$&$0$s        & \qquad\quad $0$&$001$s\\
    SVR         & $\mathit{19}$&$\mathit{5}$ & $\mathbf{0}$&$\mathbf{5}$s & $\mathbf{0}$&$\mathbf{0001}$s\\
    MLP         & $\mathbf{43}$&$\mathbf{1}$ & $200$&$0$s                 & $0$&$001$s\\
    Delaunay    & $5$&$2$                    & $1$&$0$s                   & $1$&$0$s\\
    ShepMod     & $18$&$0$                   & $\textit{0}$&$\textit{7}$s & $\mathbf{0}$&$\mathbf{0001}$s\\
    LSHEP       & $8$&$4$                    & $2$&$0$s                   & $\mathbf{0}$&$\mathbf{0001}$s\\
    Voronoi     & $0$&$5$                    & $1$&$0$s                   & $0$&$04$s\\
    BoxSplines  & $3$&$5$                    & $0$&$8$s                   & $\mathit{0}$&$\mathit{0005}$s\\
  \end{tabular}
  \caption{This average of Appendix Tables
    \ref{table:best-forest-fire}, \ref{table:best-parkinsons},
    \ref{table:best-weather}, and \ref{table:best-credit-card}
    provides a gross summary of overall results. The columns display
    (weighted equally by data set, \textit{not} points) the average
    frequency with which any algorithm provided the lowest absolute
    error approximation, the average time to fit/prepare, and the
    average time required to approximate one point. The times have
    been rounded to one significant digit, as reasonably large
    fluctuations may be observed due to implementation hardware.
    Interpolants provide the lowest error approximation for nearly one
    third of all data, while regressors occupy the other two
    thirds. This result is obtained without any customized tuning or
    preprocessing to maximize the performance of any given
    algorithm. In practice, tuning and preprocessing may have large
    effects on approximation performance.}
  \label{table:avg-performance}
\end{table}

Table \ref{table:avg-performance} summarizes results across the four
test data sets with real-valued ranges. The interpolants discussed in
this paper produce the \textit{best} approximations roughly one third
of the time, and produce competitive approximations for almost all
data sets. These test problems are almost certainly
\textit{stochastic} in nature, but the high dimension leads to greater
data sparsity and model construction cost, making interpolation more
competitive.

The major advantages to interpolation lie in the near absence of
\textit{fit} time. Delaunay, LSHEP, and ShepMod all require pairwise
distance calculations, for numerical robustness (Delaunay) and to
determine the radii of influence for data points (LSHEP and
ShepMod). At least hundreds, and sometimes hundreds of thousands of
predictions can be made by the interpolants before the most widely
used regressor (MLP) finishes fitting these relatively small data
sets.  However, the computational complexities of all interpolants
presented are greater than linear in either dimension or number of
points, whereas the regressors' nonlinear complexity in dimension
generally comes from the model fitting optimization.

The new theoretical results presented at the beginning of this chapter
directly apply to Delaunay interpolation, however the performance of
Delaunay does not appear significantly better than other algorithms on
these data sets.  This observation may be due to the stochastic nature
of the data, but it also speaks to the power of the approximations
generated by the different interpolation methods.  The strong
performance of other interpolants suggests that theoretical results
similar to those presented in this work can be achieved for the other
interpolants under reasonable assumptions.

Finally, most of the interpolants presented in this work benefit from
the ability to model \textit{any} function over real $d$-tuples with a
range that is closed under convex combinations. In general, error can
be quantified by any measure (particularly of interest may be
$L^2$, $L^\infty$, etc.). The results of the distribution prediction
case study indicate that interpolants can effectively predict
CDFs. The error analysis for that work relies on the KS statistic,
which captures the worst part of any prediction and hence provides a
conservatively large estimate of approximation error. The average
absolute errors in the predicted CDFs are always lower than the KS
statistics. However, the KS statistic was chosen as a metric because
of the important surrounding statistical theory. A nonnegligible
volume of predictions provide impressively low levels of average
absolute error in that final case study.

\section{Takeaway from Empirical Results}
\label{sec:error_conclusion}

The major contributions of this work are: 1) new uniform theoretical
error bounds for piecewise linear interpolation in arbitrary dimension;
2) an empirical evaluation across
real-world problems that demonstrates interpolants produce
competitively accurate models of multivariate phenomenon when compared
with common regressors for sparse, moderately high dimensional
problems (Section \ref{sec:error_data}); and 3) a demonstration that some
interpolants generalize to interpolation in function spaces, 
preserving monotonicity (with CDFs, e.g.), neither
of which common regressors can do.

The various interpolants discussed in this paper have been
demonstrated to effectively approximate multivariate phenomena up to
$30$ dimensions. The underlying constructions are theoretically
straightforward, interpretable, and yield reasonably accurate
predictions. Most of the interpolants' computational complexities make
them particularly suitable for applications in even higher
dimension. The major benefits of interpolation are seen when only a
small number of approximations ($\leq 1000$) are made from data and
when there are relatively few data points for the dimension (for
empirical results presented, $\log_d n \leq 5$). These findings
encourage broader application of interpolants to multivariate
approximation in science.