### 2020-08-29 3-Dissertation/Paper/thesis.tex

\documentclass[singlespace,nopageskip]{VTthesis} % nopageskip = Removes arbitrary blank pages.

%% HPC
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{mathptmx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{float}    % For controlling the placement of tables.
%% 
\usepackage{booktabs} % For formal tables
\usepackage{tikz}     % For drawing example boxes
\usetikzlibrary{positioning}
\usepackage{plain} % For plain TeX in the TOMS paper

\usepackage{hyphenat}
\hyphenation{op-tical net-works semi-conduc-tor}

%% Command for creating multiplie-line cells inside a table environment
\newcommand{\multilinecell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

%% Create a ``Lemma'' environment that does not have italics font.
\spnewtheorem{plainlemma}{Lemma}{\bf}{}
%% Create a ``Theorem'' environment that is not numbered.
\spnewtheorem*{onetheorem}{Theorem}{\bf}{}
%% Create a ``Proof'' environment that starts with 'Proof.'
\spnewtheorem*{proofdot}{Proof.}{\it}{}

% Using the following header instead will create a draft copy of your thesis
%\documentclass[doublespace,draft]{VTthesis}

% Title of your thesis
\title{Interpolants, Error Bounds, and Mathematical Software for Modeling and Predicting Variability in Computer Systems}

% You should include 3-5 keywords, separated by commas
\keywords{Approximation Theory, Numerical Analysis, High Performance Computing, Computer Security, Nonparametric Statistics, Mathematical Software}

% Your name, including middle initial(s)
\author{Thomas C. H. Lux}

% Change this to your program, e.g. Physics, Civil Engineering, etc.
\program{Computer Science}

% Change this to your degree, e.g. Master of Science, Master of Art, etc.
\degree{Doctor of Philosophy}

% This should be your defense date:
\submitdate{August 14, 2020}

% Committee members. Only have five readers and one chair available.
% Only use the ones you need and don't include the ones you don't need.
% You can also declare a Co-advisor. If you do, the principal and co-advisors
% will be listed as co-advisors on the title page.  Per the VT ETD standards, 
% you should not include titles or educational qualifications such as PhD or Dr.
% You should, however, include middle initials if possible.
\principaladvisor{Layne Watson}
\coadvisor{Yili Hong}
\firstreader{Kirk Cameron}
\secondreader{Bert Huang}
\thirdreader{Gang Wang}
\fourthreader{Yang Cao}

% The dedication and acknowledgement pages are optional. Comment them out to remove them.
\dedication{
For Alex Faustie Grant, my dream teammate.
}
\acknowledge{
Finishing a Ph.D. is hard work. It's not always fun. And sometimes it's stressful. But the people I've met along the way, the things I've learned, and the people from all paths of life that have supported me throughout this journey made all the difference. My gratitudes are endless, but I'll do my best to enumerate some here.
{\vspace{2mm} \\}
First and foremost I want to thank my advisor Dr. Watson for his enduring push for deep scientific rigor and always reminding me of my most important job, to ``think!'' While reading that without context might fall short, Dr. Watson's incredible effort, earnest care, and impressive repertoire of knowledge afforded me many great lessons throughout my graduate tenure that I value dearly. His genuine feedback has helped me become a better researcher and he's someone I won't hesitate to consult in the future.
{\vspace{2mm} \\}
I want to thank my coadvisor Dr. Hong for deepening my knowledge of statistics and introducing me to countless relevant and interesting lines of research. His support and direction drove the statistical significance of my work. In addition I want to thank all of my committee members for their care and contributions, Dr. Cameron for asking important practical questions and encouraging a bigger picture cohesion throughout my work, Dr. Huang and Dr. Wang for ensuring theoretical groundedness and relevance to modern research as well as being great teachers, and Dr. Cao for motivating many connections between my work and other areas of applied computational science.
{\vspace{-1mm} \\}
I was uniquely lucky in that the entirety of my work at Virginia Tech was supported by the VarSys project. Those team members, both students and faculty, guided my growth as a researcher and my research direction as a whole. I thank them all for the many conversations about fascinating and diverse topics in computer science, as well as constant feedback and collaboration that broadened my interests and drove me to produce higher quality results.
{\vspace{2mm} \\}
Where would I be without my closest friends through graduate school. I sincerely thank all of Grad Council, those who helped found the group and see it to its great success as a leading computer science graduate council in the nation. I am especially grateful for my lab mates Tyler Chang and Ayaan Kazerouni for the lunches around Blacksburg, encouraging conversations, and general support that made graduate life fun and memorable in ways that it would never have been otherwise.
{\vspace{2mm} \\}
Another thanks goes to my family for their continued love and support, particularly my parents Karen and John Lux, brothers Dylan and Austen, sisters Amy, and Arial, and grandmothers Gooma and Mor Mor. They have made all the difference and I am unbelievably grateful for them all.
{\vspace{2mm} \\}
Lastly I would like to thank my fianc\'e Alex Grant for her enduring love and support over the last four years. I couldn't imagine getting this far without our fun trips to Kentucky, summers together, and Skype dinner and movie dates that made the distance between us from Blacksburg to St. Louis bearable.
{\vspace{10mm} \\}
This work was supported in part by the National Science Foundation under Grant Numbers CNS-1565314 and CNS-1838271. Any opinions, findings, and conclusions or recommendations expressed in this material are my own and do not necessarily reflect the views of the National Science Foundation.
}

% The abstract is required and should be <=250 words for thesis, <=350 words for dissertation.
\abstract{Function approximation is an important problem. This work presents applications of interpolants to modeling random variables. Specifically, this work studies the prediction of distributions of random variables applied to computer system throughput variability. Existing approximation methods including multivariate adaptive regression splines, support vector regressors, multilayer perceptrons, Shepard variants, and the Delaunay mesh are investigated in the context of computer variability modeling. New methods of approximation using Box splines, Voronoi cells, and Delaunay for interpolating distributions of data with moderately high dimension are presented and compared with existing approaches. Novel theoretical error bounds are constructed for piecewise linear interpolants over functions with a Lipschitz continuous gradient. Finally, a mathematical software that constructs monotone quintic spline interpolants for distribution approximation from data samples is proposed.}

% The general audience abstract is required. There are currently no word limits.
\abstractgenaud{It is common for scientists to collect data on something they are studying. Often scientists want to create a (predictive) model of that phenomenon based on the data, but the choice of \textit{how} to model the data is a difficult one to answer. This work proposes methods for modeling data that operate under very few assumptions that are broadly applicable across science. Finally, a software package is proposed that would allow scientists to better understand the \textit{true} distribution of their data given relatively few observations.}

\begin{document}
% The following lines set up the front matter of your thesis or dissertation and is required to ensure proper formatting per the VT ETD standards. 
\frontmatter
\maketitle
\tableofcontents

% The list of figures and tables are now optional per the official ETD standards.  Unless you have a very good reason for removing them, you should leave these lists in the document. Comment them out to remove them.
\listoffigures
\listoftables

%% \printnomenclature %creates a list of abbreviations. Comment out to remove it. 
%% ABBR is described here.
%% \nomenclature{ABBR}{abbreviation}

% The following sets up the document for the main part of the thesis or dissertation. Do not comment out or remove this line.
\mainmatter

%% Introduction of variability modeling, the importance of modeling in science in general, then dive into the specific application of modeling techniques to systems (and security).
\chapter{The Importance and Applications of Variability} \label{ch:apps}
\input{1-apps}

%% Describe all of the interpolants and regressors that will be used throughout this work. (except meshes)
\chapter{Algorithms for Constructing Approximations} \label{ch:algs}
\input{2-algs}

%% Demonstrate the naive application of interpolants and regressors to data. Discuss results.
\chapter{Na\"{\i}ve Approximations of Variability} \label{ch:naive}
\input{3-naive}

%% Introduce the theory on box splines. Introduce the algorithms that rely on box splines. Show the results of the code that produces box splines.
%% Demonstrate the application of these meshes (and box splines particularly) to approximation problems. Discuss results.
\chapter{Box Splines: Uses, Constructions, and Applications} \label{ch:boxes}
\input{4-boxes}

%% Introduce a measurement for error when predicting distribution outputs, then show what happens when system behavior is predicted as a distribution.
\chapter{Stronger Approximations of Variability} \label{ch:strong}
\input{5-strong}

%% Construct error theory that works for all observed applications to this point.
\chapter{An Error Bound on Piecewise Linear Interpolation} \label{ch:error}
%% Apply approximation algorithms to a final set of problems. Discuss results.
\input{6-error}

%% Propose new method to improve predictions for distributions.
%% \chapter{Improving Variability Estimates with Monotone $C^2$ Splines} 
\chapter{A Package for Monotone Quintic Spline Interpolation} \label{ch:splines}
\input{7-splines}

%% Lessons learned, and future directions for research
\chapter{Conclusions on Modeling and Approximating Variability in Computer Systems} \label{ch:concs}
\input{8-conclusions}

%% Bibliography.
\bibliography{thesis}
\bibliographystyle{plainnat}

% This formats the chapter name to appendix to properly define the headers:
\appendix

% Add your appendices here. You must leave the appendices enclosed in the appendices environment in order for the table of contents to be correct.
\begin{appendices}
\chapter{Error Bound Empirical Test Results} \label{app:error}

\input{9-error-app}

\chapter{\texttt{MQSI} Package} \label{app:mqsi}

%% -------------------------------------------------------------------
%%     Setup for including a code listing in a paper.

%% Allow TT fonts to be scaled with LaTeX macros like '\tiny'.
\newcommand{\ttscale}{s*[0.8]}
\DeclareFontFamily{OT1}{cmtt}{\hyphenchar\font -1 }
\DeclareFontShape{OT1}{cmtt}{m}{n}{
        <-9>   \ttscale cmtt8
        <9-10> \ttscale cmtt9
        <10-12>\ttscale cmtt10
        <12->  \ttscale cmtt12
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{bf}{
        <->    \ttscale cmbftt10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{it}{
        <->    \ttscale cmitt10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{sl}{
        <->    \ttscale cmsltt10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{sc}{
        <->    \ttscale cmtcsc10
      }{}
\DeclareFontShape{OT1}{cmtt}{m}{ui}
       {<->ssub*cmtt/m/it}{}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
       {<->ssub*cmtt/m/n}{}
\DeclareFontShape{OT1}{cmtt}{bx}{it}
       {<->ssub*cmtt/m/it}{}
\DeclareFontShape{OT1}{cmtt}{bx}{ui}
       {<->ssub*cmtt/m/it}{}

%% Define a new color that is subtle but can be used for keywords.
\definecolor{bluegray}{rgb}{0.2, 0.3, 0.5}

\lstset{% general command to set parameter(s)
  frame=,                                  % disable the frame lines
  basicstyle=\color{black}\small\ttfamily, % whole listing style
  keywordstyle=\color{bluegray},           % language keyword color
  identifierstyle=,                        % no identifier
  commentstyle=\color{gray},               % comment color
  showstringspaces=false,                  % no special string spaces
  numbers=none                             % number lines
}
  %% numberstyle=\tiny,                       % make style of numbers tiny
  %% stepnumber=3,                            % show numbers every * lines
  %% numbersep=5pt%

%% -------------------------------------------------------------------

\lstinputlisting{MQSI/README}
\lstinputlisting[language=Fortran]{MQSI/REAL_PRECISION.f90}
\lstinputlisting[language=Fortran]{MQSI/MQSI.f90}
\lstinputlisting[language=Fortran]{MQSI/SPLINE.f90}
\lstinputlisting[language=Fortran]{MQSI/EVAL_BSPLINE.f90}

\chapter{Box Spline Subroutine} \label{app:boxspline}

\lstinputlisting[language=Fortran]{boxspline.f90}

\end{appendices}

\end{document}

%****************************************************************************
% Below are some general suggestions for writing your dissertation:
%
% 1. Label everything with a meaningful prefix so that you
%    can refer back to sections, tables, figures, equations, etc.
%    Usage \label{<prefix>:<label_name>} where some suggested
%    prefixes are:
%			ch: Chapter
%     		se: Section
%     		ss: Subsection
%     		sss: Sub-subsection
%			app: Appendix
%     		ase: Appendix section
%     		tab: Tables
%     		fig: Figures
%     		sfig: Sub-figures
%     		eq: Equations
%
% 2. The VTthesis class provides for natbib citations. You should upload
%	 one or more *.bib bibtex files. Suppose you have two bib files: some_refs.bib and 
%    other_refs.bib.  Then your bibliography line to include them
%    will be:
%      \bibliography{some_refs, other_refs}
%    where multiple files are separated by commas. In the body of 
%    your work, you can cite your references using natbib citations.
%    Examples:
%      Citation                     Output
%      -------------------------------------------------------
%      \cite{doe_title_2016}        [18]
%      \citet{doe_title_2016}       Doe et al. [18]
%      \citet*{doe_title_2016}      Doe, Jones, and Smith [18]
%
%    For a complete list of options, see
%      https://www.ctan.org/pkg/natbib?lang=en
%
% 3. Here is a sample table. Notice that the caption is centered at the top. Also
%    notice that we use booktabs formatting. You should not use vertical lines
%    in your tables.
% 
%				\begin{table}[htb]
%					\centering
%					\caption{Approximate computation times in hh:mm:ss for full order 						versus reduced order models.}
%					\begin{tabular}{ccc}
%						\toprule
%						& \multicolumn{2}{c}{Computation Time}\\
%						\cmidrule(r){2-3}
%						$\overline{U}_{in}$ m/s & Full Model & ROM \\
%						\midrule
%						0.90 & 2:00:00 & 2:08:00\\
%						0.88 & 2:00:00 & 0:00:03\\
%						0.92 & 2:00:00 & 0:00:03\\
%						\midrule
%						Total & 6:00:00 & 2:08:06\\
%						\bottomrule
%					\end{tabular}
%					\label{tab:time_rom}
%				\end{table}
% 
% 4. Below are some sample figures. Notice the caption is centered below the
%    figure.
%    a. Single centered figure:
%					\begin{figure}[htb]
%						\centering
%						\includegraphics[scale=0.5]{my_figure.eps}
%						\caption{Average outlet velocity magnitude given an average  
%				        input velocity magnitude of 0.88 m/s.} 
%						\label{fig:output_rom}
%					\end{figure}
%    b. Two by two grid of figures with subcaptions
%					\begin{figure}[htb]
%						\centering
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_1_1.eps}
%							\caption{Subcaption number one}
%							\label{sfig:first_subfig}
%						\end{subfigure}
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_1_2.png}
%							\caption{Subcaption number two}
%							\label{sfig:second_subfig}
%						\end{subfigure}
%
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_2_1.pdf}
%							\caption{Subcaption number three}
%							\label{sfig:third_subfig}
%						\end{subfigure}
%						\begin{subfigure}[h]{0.45\textwidth}
%							\centering
%							\includegraphics[scale=0.4]{figure_2_2.eps}
%							\caption{Subcaption number four}
%							\label{sfig:fourth_subfig}
%						\end{subfigure}
%						\caption{Here is my main caption describing the relationship 							between the 4 subimages}
%						\label{fig:main_figure}
%					\end{figure}
%
%----------------------------------------------------------------------------
%
% The following is a list of definitions and packages provided by VTthesis:
%
% A. The following packages are provided by the VTthesis class:
%      amsmath, amsthm, amssymb, enumerate, natbib, hyperref, graphicx, 
%      tikz (with shapes and arrows libraries), caption, subcaption,
%      listings, verbatim
%
% B. The following theorem environments are defined by VTthesis:
%      theorem, proposition, lemma, corollary, conjecture
% 
% C. The following definition environments are defined by VTthesis:
%      definition, example, remark, algorithm
%
%----------------------------------------------------------------------------
%
%  I hope this template file and the VTthesis class will keep you from having 
%  to worry about the formatting and allow you to focus on the actual writing.
%  Good luck, and happy writing.
%    Alan Lattimer, VT, 2016
%
%****************************************************************************